
---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bonsaiforest2 🌳

The goal of `bonsaiforest2` is to **simplify fitting and interpreting Bayesian models for subgroup analysis** in clinical trials. It leverages the power of the `brms` package to:

  * Distinguish between **prognostic** (baseline predictors) and **predictive** (treatment effect modifiers) factors.
  * Apply differential **shrinkage priors** (like Horseshoe or R2D2) to explore subgroup effects robustly.
  * Calculate interpretable **marginal treatment effects** using a counterfactual approach.
  * Generate publication-ready **forest plots**.

## Installation

You can install the development version of `bonsaiforest2` from its GitLab repository:

```{r, eval=FALSE}
# install.packages("remotes") 
remotes::install_github("openpharma/bonsaiforest2")
```

## Example

This is a basic example showing the main workflow: fitting a simple continuous outcome model with shrinkage on a subgroup interaction.

```{r example, message=FALSE, warning=FALSE, results='hide'}
library(bonsaiforest2)
library(brms) # Needed for backend

# 1. Minimal data setup
set.seed(123)
n <- 50 # Small n for README example
sim_data_readme <- data.frame(
  outcome = rnorm(n),
  trt = factor(sample(0:1, n, replace = TRUE)),
  age = rnorm(n, 50, 10),
  region = factor(sample(c("A", "B"), n, replace = TRUE))
)

# 2. Fit a simple model (use very few iterations for speed!)
# Shrink the treatment:region interaction
fit <- run_brms_analysis(
  data = sim_data_readme,
  response_formula_str = "outcome ~ trt",
  response_type = "continuous",
  unshrunk_prognostic_formula_str = "~ age",     # Adjust for age
  shrunk_predictive_formula_str = "~ trt:region", # Explore region interaction
  chains = 1, iter = 50, warmup = 25, refresh = 0, # Keep it FAST
  backend = "cmdstanr" # Optional: Specify backend if needed
)

# 3. Summarize marginal effects (will auto-detect 'region')
effect_summary <- summary_subgroup_effects(
  brms_fit = fit,
  original_data = sim_data_readme,
  trt_var = "trt",
  response_type = "continuous"
)

# 4. Plot the results
plot(effect_summary) # Display plot
```

**Summarized Effects:**

```{r show-summary, echo=FALSE}
# Display the summary table generated above
print(effect_summary$estimates)
```

