---
title: "Forest Plot: Comparing Treatment Effect Estimators Against Truth"
subtitle: "Random Simulation from Scenario 3"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 12
    fig_height: 8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = FALSE,
  fig.width = 14,
  fig.height = 10
)
```

# Overview

This document demonstrates how to:
1. Load a randomly selected simulation from Scenario 3
2. Fit multiple treatment effect estimators (Naive Subgroup and Population)
3. Compare estimates against the ground truth
4. Visualize results in a forest plot

---

# 1. Load Libraries and Data

```{r load_libraries}
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)
library(tibble)
library(purrr)

# Set working directory to Continuous folder
# Try to use the document directory if available, otherwise use explicit path
tryCatch({
  setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
}, error = function(e) {
  # Fallback: set to Continuous folder explicitly
  setwd("~/bonsaiforest2/simulations/paper/Continuous")
})

# Verify we're in the right directory
cat("Working directory:", getwd(), "\n")
```

```{r load_truth_and_scenarios}
# Load ground truth
load("Scenarios/truth.RData")

# Load scenario 3 data (list of 1000 simulations)
scenario3_all <- readRDS("Scenarios/scenario3.rds")

# Randomly select one simulation
set.seed(0)  # Change seed for different random selections
random_sim_id <- sample(1:length(scenario3_all), 1)
sim_data <- as.data.frame(scenario3_all[[random_sim_id]])

cat(sprintf("Selected simulation #%d from Scenario 3\n", random_sim_id))
cat(sprintf("Sample size: %d\n", nrow(sim_data)))
cat(sprintf("Covariates: %s\n", paste(names(sim_data)[!names(sim_data) %in% c("trt", "Y")], collapse = ", ")))

# Get truth data for scenario 3
truth_scen3 <- simulation_truth %>%
  filter(scenario_no == 3)

head(truth_scen3)
```

---

# 2. Naive Subgroup Analysis

Fit a linear model `Y ~ trt` for each subgroup, using only data from that subgroup level.

```{r subgroup_analysis}
# Define subgroup variables
subgroup_vars <- c("X1", "X2", "X4", "X8", "X11cat", "X14cat", "X17cat")

# Create unique combinations of (subgroup_var, level)
subgroup_combinations <- expand.grid(
  subgroup_var = subgroup_vars,
  stringsAsFactors = FALSE
) %>%
  as_tibble() %>%
  rowwise() %>%
  mutate(
    levels_list = list(unique(na.omit(sim_data[[subgroup_var]])))
  ) %>%
  unnest(levels_list) %>%
  rename(subgroup_level = levels_list) %>%
  mutate(
    subgroup_id = paste0(subgroup_var, ".", as.character(subgroup_level))
  ) %>%
  ungroup()

cat(sprintf("Analyzing %d unique subgroup combinations\n", nrow(subgroup_combinations)))

# Fit a model for each unique subgroup combination
subgroup_results <- map_df(seq_len(nrow(subgroup_combinations)), function(i) {
  
  row <- subgroup_combinations[i, ]
  subgroup_var <- row$subgroup_var
  subgroup_level <- row$subgroup_level
  subgroup_id <- row$subgroup_id
  
  # Subset data to this subgroup (convert to character to handle factors safely)
  df_sub <- sim_data %>% 
    filter(!is.na(.data[[subgroup_var]]), as.character(.data[[subgroup_var]]) == as.character(subgroup_level))
  
  n_obs <- nrow(df_sub)
  
  if (n_obs < 2) {
    return(tibble(
      subgroup_id = subgroup_id,
      subgroup_var = subgroup_var,
      subgroup_level = as.character(subgroup_level),
      estimator = "Naive Subgroup",
      n = n_obs,
      n_trt0 = NA_real_,
      n_trt1 = NA_real_,
      trt_effect = NA_real_,
      trt_se = NA_real_,
      trt_t_stat = NA_real_,
      trt_p_value = NA_real_,
      ci_lower = NA_real_,
      ci_upper = NA_real_,
      mean_y_trt0 = NA_real_,
      mean_y_trt1 = NA_real_,
      sd_y_trt0 = NA_real_,
      sd_y_trt1 = NA_real_
    ))
  }
  
  # Fit linear model: Y ~ trt
  fit <- lm(Y ~ trt, data = df_sub)
  tidy_res <- broom::tidy(fit) %>% filter(term == "trt")
  
  # Get counts and means by treatment group
  summary_stats <- df_sub %>%
    group_by(trt) %>%
    summarise(
      n_trt = n(),
      mean_y = mean(Y),
      sd_y = sd(Y),
      .groups = "drop"
    ) %>%
    pivot_wider(
      names_from = trt,
      values_from = c(n_trt, mean_y, sd_y),
      names_glue = "{.value}_trt{trt}",
      values_fill = NA_real_
    )
  
  # Safely extract values (even if treatment group is missing)
  tibble(
    subgroup_id = subgroup_id,
    subgroup_var = subgroup_var,
    subgroup_level = as.character(subgroup_level),
    estimator = "Naive Subgroup",
    n = n_obs,
    n_trt0 = ifelse("n_trt_trt0" %in% names(summary_stats), summary_stats$n_trt_trt0, NA_real_),
    n_trt1 = ifelse("n_trt_trt1" %in% names(summary_stats), summary_stats$n_trt_trt1, NA_real_),
    trt_effect = tidy_res$estimate[1],
    trt_se = tidy_res$std.error[1],
    trt_t_stat = tidy_res$statistic[1],
    trt_p_value = tidy_res$p.value[1],
    ci_lower = trt_effect - 1.96 * trt_se,
    ci_upper = trt_effect + 1.96 * trt_se,
    mean_y_trt0 = ifelse("mean_y_trt0" %in% names(summary_stats), summary_stats$mean_y_trt0, NA_real_),
    mean_y_trt1 = ifelse("mean_y_trt1" %in% names(summary_stats), summary_stats$mean_y_trt1, NA_real_),
    sd_y_trt0 = ifelse("sd_y_trt0" %in% names(summary_stats), summary_stats$sd_y_trt0, NA_real_),
    sd_y_trt1 = ifelse("sd_y_trt1" %in% names(summary_stats), summary_stats$sd_y_trt1, NA_real_)
  )
})

cat(sprintf("Generated %d subgroup estimates\n", nrow(subgroup_results)))
head(subgroup_results)
```

---

# 3. Naive Population Analysis

Fit a linear model `Y ~ trt` using all data (overall/global effect).

```{r population_analysis}
# Fit population-level model
pop_fit <- lm(Y ~ trt, data = sim_data)
pop_coef <- broom::tidy(pop_fit) %>% filter(term == "trt")

# Get summary statistics
pop_summary <- sim_data %>%
  group_by(trt) %>%
  summarise(
    n = n(),
    mean_y = mean(Y),
    sd_y = sd(Y),
    .groups = "drop"
  )

# Extract values for population results
pop_trt_effect <- pop_coef$estimate[1]
pop_trt_se <- pop_coef$std.error[1]
pop_ci_lower <- pop_trt_effect - 1.96 * pop_trt_se
pop_ci_upper <- pop_trt_effect + 1.96 * pop_trt_se

population_results <- tibble(
  subgroup_id = "Overall",
  subgroup_var = "Overall",
  subgroup_level = "All",
  estimator = "Naive Population",
  n = nrow(sim_data),
  n_trt0 = pop_summary$n[1],
  n_trt1 = pop_summary$n[2],
  trt_effect = pop_trt_effect,
  trt_se = pop_trt_se,
  ci_lower = pop_ci_lower,
  ci_upper = pop_ci_upper,
  mean_y_trt0 = pop_summary$mean_y[1],
  mean_y_trt1 = pop_summary$mean_y[2],
  sd_y_trt0 = pop_summary$sd_y[1],
  sd_y_trt1 = pop_summary$sd_y[2]
)

cat("Population-level estimates:\n")
print(population_results[, c("estimator", "trt_effect", "trt_se", "ci_lower", "ci_upper")])
```

---

# 4. Merge with Ground Truth

```{r merge_with_truth}
# Check structure of truth data
cat("Columns in truth_scen3: ", paste(names(truth_scen3), collapse = ", "), "\n")
str(truth_scen3)

# Prepare truth data for merging
truth_for_merge <- truth_scen3 %>%
  mutate(
    subgroup_level_char = as.character(level),
    subgroup_id = paste0(subgroup_var, ".", subgroup_level_char),
    truth_value = trt_effect
  )

# Calculate overall truth (mean across all subgroups)
overall_truth <- truth_scen3$trt_effect %>% mean()

truth_pop <- tibble(
  subgroup_id = "Overall",
  subgroup_var = "Overall",
  truth_value = overall_truth
)

truth_all <- bind_rows(
  truth_for_merge %>% dplyr::select(subgroup_id, subgroup_var, truth_value),
  truth_pop
)

# Combine all estimates
all_estimates <- bind_rows(subgroup_results, population_results)

# Merge with truth
comparison_data <- all_estimates %>%
  left_join(
    truth_all %>% dplyr::select(subgroup_id, truth_value),
    by = "subgroup_id"
  ) %>%
  mutate(
    bias = trt_effect - truth_value,
    coverage = (ci_lower <= truth_value) & (truth_value <= ci_upper)
  )

head(comparison_data, 15)
```

---

# 5. Summary Table

```{r summary_table}
summary_table <- comparison_data %>%
  group_by(estimator, subgroup_var) %>%
  summarise(
    n_subgroups = n(),
    mean_estimate = mean(trt_effect, na.rm = TRUE),
    mean_bias = mean(bias, na.rm = TRUE),
    mean_se = mean(trt_se, na.rm = TRUE),
    coverage = mean(coverage, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(subgroup_var, estimator)

knitr::kable(summary_table, caption = "Summary of Estimates by Estimator and Subgroup Variable")
```

---

# 6. Forest Plot: All Subgroups

```{r forest_plot_all}
# Prepare data for forest plot
forest_data <- comparison_data %>%
  filter(estimator == "Naive Subgroup") %>%
  arrange(subgroup_var, subgroup_level) %>%
  mutate(
    group_label = paste0(subgroup_var, " = ", subgroup_level),
    group_order = row_number()
  )

# Create forest plot
p_forest <- ggplot(forest_data, aes(x = reorder(group_label, -group_order), y = trt_effect)) +
  geom_point(aes(color = "Estimate"), size = 3, position = position_dodge(width = 0.2)) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper, color = "Estimate"),
    width = 0.3,
    position = position_dodge(width = 0.2)
  ) +
  geom_point(
    aes(y = truth_value, color = "Truth"),
    size = 3,
    shape = 4,
    stroke = 2,
    position = position_dodge(width = 0.2)
  ) +
  facet_wrap(~subgroup_var, scales = "free", ncol = 2) +
  coord_flip() +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    strip.text = element_text(face = "bold", size = 11),
    legend.position = "bottom"
  ) +
  labs(
    title = "Forest Plot: Naive Subgroup Estimates vs. Ground Truth",
    subtitle = sprintf("Scenario 3, Simulation #%d", random_sim_id),
    x = "Subgroup",
    y = "Treatment Effect Estimate (95% CI)",
    color = "Type"
  ) +
  scale_color_manual(
    values = c("Estimate" = "steelblue", "Truth" = "red")
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", alpha = 0.5)

print(p_forest)
```

---

# 7. Forest Plot: Population vs. Subgroup Estimates

```{r forest_plot_comparison}
# Select representative subgroups for cleaner visualization
representative_subgroups <- comparison_data %>%
  filter(estimator == "Naive Subgroup") %>%
  group_by(subgroup_var) %>%
  slice(1:3) %>%  # First 3 levels of each variable
  ungroup() %>%
  dplyr::select(subgroup_id)

# Add population estimate
comparison_data_subset <- comparison_data %>%
  filter(subgroup_id %in% c(representative_subgroups$subgroup_id, "Overall")) %>%
  mutate(
    group_label = ifelse(subgroup_id == "Overall", "Overall Population", subgroup_id),
    estimator = factor(estimator, levels = c("Naive Population", "Naive Subgroup"))
  )

# Forest plot comparing all estimators
comparison_data_subset <- comparison_data_subset %>%
  mutate(
    group_order = row_number()
  )

p_forest_comp <- ggplot(comparison_data_subset, aes(x = reorder(group_label, group_order), y = trt_effect)) +
  geom_point(aes(color = estimator), size = 3, position = position_dodge(width = 0.5)) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper, color = estimator),
    width = 0.4,
    position = position_dodge(width = 0.5)
  ) +
  geom_point(
    aes(y = truth_value),
    color = "red",
    size = 3.5,
    shape = 4,
    stroke = 2,
    position = position_dodge(width = 0.5)
  ) +
  coord_flip() +
  theme_minimal() +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "top"
  ) +
  labs(
    title = "Forest Plot: Population vs. Subgroup Estimates",
    subtitle = sprintf("Scenario 3, Simulation #%d (Red ✕ = Ground Truth)", random_sim_id),
    x = "Subgroup",
    y = "Treatment Effect Estimate (95% CI)",
    color = "Estimator"
  ) +
  scale_color_manual(
    values = c("Naive Population" = "steelblue", "Naive Subgroup" = "orange")
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", alpha = 0.5)

print(p_forest_comp)
```

---

# 8. Performance Metrics

```{r performance_metrics}
performance <- comparison_data %>%
  group_by(estimator) %>%
  summarise(
    n_estimates = n(),
    mean_bias = mean(bias, na.rm = TRUE),
    rmse_bias = sqrt(mean(bias^2, na.rm = TRUE)),
    mean_se = mean(trt_se, na.rm = TRUE),
    empirical_coverage = mean(coverage, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    mean_bias = round(mean_bias, 4),
    rmse_bias = round(rmse_bias, 4),
    mean_se = round(mean_se, 4),
    empirical_coverage = round(empirical_coverage, 4)
  )

knitr::kable(
  performance,
  caption = "Performance Metrics by Estimator",
  col.names = c("Estimator", "N Estimates", "Mean Bias", "RMSE Bias", "Mean SE", "Coverage Rate")
)
```

---

# 9. Session Info

```{r session_info}
sessionInfo()
```
