---
title: "Paper Simulation Results Analysis"
subtitle: "TTE and Continuous Outcomes - Selected Figures and Tables"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.width = 14, fig.height = 9)

library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(ggtext)
library(knitr)
library(kableExtra)
library(rlang)
```

---

# TTE Simulation Results for Paper

---

## Load TTE Results

```{r load_tte_results, include=FALSE}
# Set paths for TTE
tte_base_path <- "/home/pedreram/bonsaiforest2/simulations/paper/TTE"
tte_data_file <- file.path(tte_base_path, "tte_results_prepared.rds")

# Check if prepared data exists
if (!file.exists(tte_data_file)) {
  cat("ERROR: TTE prepared data file not found!\n")
  cat("Please run: Rscript", file.path(tte_base_path, "prepare_tte_results.R\n"))
  stop("Missing prepared TTE data")
}

# Load merged TTE results (with estimates and truth)
tte_all_results <- readRDS(tte_data_file) 

cat("✓ Loaded TTE merged results from:", tte_data_file, "\n")
cat("  Rows:", nrow(tte_all_results), "\n")
cat("  Estimators:", n_distinct(tte_all_results$estimator), "\n")
cat("  Scenarios:", unique(sort(tte_all_results$scenario_no)), "\n\n")
```

```{r tte_calculate_metrics}
# Calculate RMSE, Bias, and Coverage directly from individual replications
# RMSE computed across ALL replications and subgroups combined

# First: calculate per-subgroup metrics for detailed analysis
# IMPORTANT: Compute signed bias first, then take absolute value (like TTE_Results_Analysis.Rmd)
tte_subgroup_metrics <- tte_all_results %>%
  mutate(
    error = estimate_log - truth_log,
    ci_contains_truth = (truth_log >= ci_lower) & (truth_log <= ci_upper)
  ) %>%
  group_by(scenario_no, estimator, join_key) %>%
  summarise(
    n_replications = n(),
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    bias_log = mean(error, na.rm = TRUE),  # Signed bias per subgroup
    abs_bias_log = abs(bias_log),  # Take absolute value of signed bias (not mean of abs)
    coverage = mean(ci_contains_truth, na.rm = TRUE),
    truth_log = first(truth_log),
    truth_ahr = first(truth_ahr),
    .groups = 'drop'
  )

# Second: calculate overall metrics per estimator-scenario
# RMSE is computed from ALL replication-subgroup pairs combined
tte_performance_by_scenario <- tte_all_results %>%
  mutate(
    error = estimate_log - truth_log,
    ci_contains_truth = (truth_log >= ci_lower) & (truth_log <= ci_upper)
  ) %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    n_subgroups = n_distinct(join_key),
    n_total_pairs = n(),
    # RMSE: sqrt(mean of all squared errors across all reps and subgroups)
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    Bias = mean(error, na.rm = TRUE),
    AbsBias = mean(abs(error), na.rm = TRUE),
    Coverage = mean(ci_contains_truth, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(as.numeric(scenario_no), estimator)

# Separate datasets: main (exclude Global HN) and supplementary (Global HN only)
tte_is_global_hn <- grepl("Global_HN", tte_performance_by_scenario$estimator)
tte_performance_main <- tte_performance_by_scenario[!tte_is_global_hn, ]
tte_performance_supplementary <- tte_performance_by_scenario[tte_is_global_hn, ]

# Overall metrics across all scenarios
tte_performance_overall <- tte_all_results %>%
  mutate(
    error = estimate_log - truth_log,
    ci_contains_truth = (truth_log >= ci_lower) & (truth_log <= ci_upper)
  ) %>%
  group_by(estimator) %>%
  summarise(
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    Bias = mean(error, na.rm = TRUE),
    AbsBias = mean(abs(error), na.rm = TRUE),
    Coverage = mean(ci_contains_truth, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(estimator)

cat("TTE performance metrics calculated.\n")
```

---

## TTE Helper Functions

```{r tte_helpers}
# Beautify estimator names for TTE
beautify_tte_estimator <- function(estimator_name) {
  dplyr::case_when(
    estimator_name == "population" ~ "Population",
    estimator_name == "subgroup" ~ "Subgroup",
    # Global RHS
    grepl("Global_RHS.*theta0_1.*s_2", estimator_name) ~ "Global - RHS(τ₀=1, s=2)",
    grepl("Global_RHS.*theta0_delta_plan_half.*s_2", estimator_name) ~ "Global - RHS(τ₀=δ<sub>plan</sub>/2, s=2)",
    grepl("Global_RHS.*theta0_delta_plan.*s_2", estimator_name) ~ "Global - RHS(τ₀=δ<sub>plan</sub>, s=2)",
        grepl("Global_RHS.*theta0_delta_plan_half_10.*s_2", estimator_name) ~ "Global - RHS(τ₀=δ<sub>plan</sub>/10, s=2)",
    # OVAT HN
    grepl("OVAT_HN.*phi_1$", estimator_name) ~ "OVAT - HN(φ=1)",
    grepl("OVAT_HN.*phi_delta_plan_half", estimator_name) ~ "OVAT - HN(φ=δ<sub>plan</sub>/2)",
    grepl("OVAT_HN.*phi_delta_plan$", estimator_name) ~ "OVAT - HN(φ=δ<sub>plan</sub>)",
    TRUE ~ estimator_name
  )
}

# Beautify scenario names for TTE
beautify_tte_scenario <- function(scenario_no) {
  dplyr::case_when(
    scenario_no == "1" ~ "Scenario 1:\nPositive (homogeneous)",
    scenario_no == "2" ~ "Scenario 2:\nPositive (except 1)",
    scenario_no == "3" ~ "Scenario 3:\nMild heterogeneity",
    scenario_no == "4" ~ "Scenario 4:\nStrong heterogeneity",
    TRUE ~ paste("Scenario", scenario_no)
  )
}

# Format function
display <- function(x, digits = 3) {
  formatC(x, format = "f", digits = digits)
}

# Percentage function
getPercentage <- function(x, digits = 1) {
  val <- 100 * x
  ifelse(is.na(val), "NA",
    paste(formatC(val, digits = digits, format = "f"), "%", sep = "")
  )
}
```

---

## TTE Figure: Standardized Overall RMSE

```{r tte_figure_rmse, fig.width=12, fig.height=6}
# Prepare data: standardize RMSE within each scenario by baseline (subgroup estimator)
# Filter to main section (exclude Global HN)
tte_plot_data <- tte_performance_main %>%
  mutate(
    scenario_no_numeric = as.numeric(scenario_no),
    estimator_display = beautify_tte_estimator(estimator),
    scenario_label = beautify_tte_scenario(scenario_no)
  )

# Use subgroup estimator as baseline for standardization
tte_baselines <- tte_performance_by_scenario %>%
  filter(estimator == "subgroup") %>%
  dplyr::select(scenario_no, baseline_rmse = RMSE)

tte_plot_data <- tte_plot_data %>%
  mutate(estimator_display = ifelse(estimator=="Global_RHS_theta0_delta_plan_half_10_s_2", "Global - RHS(τ₀=δ<sub>plan</sub>/10, s=2)",
                                    estimator_display))


# Standardize RMSE
tte_plot_data_standardized <- tte_plot_data %>%
  left_join(tte_baselines, by = "scenario_no") %>%
  mutate(
    RMSE_standardized = RMSE / baseline_rmse
  )

# Create plot with manual color scheme
tte_color_map <- c(
  "Subgroup" = "red",
  "Population" = "orange",
  "OVAT - HN(φ=1)" = "skyblue",
  "OVAT - HN(φ=δ<sub>plan</sub>/2)" = "royalblue",
  "OVAT - HN(φ=δ<sub>plan</sub>)" = "navy",
  "Global - RHS(τ₀=1, s=2)" = "palegreen1",
  "Global - RHS(τ₀=δ<sub>plan</sub>, s=2)" = "palegreen3",
  "Global - RHS(τ₀=δ<sub>plan</sub>/2, s=2)" = "forestgreen",
  "Global - RHS(τ₀=δ<sub>plan</sub>/10, s=2)" = "darkgreen"
)

ggplot(tte_plot_data_standardized, aes(x = scenario_label, y = RMSE_standardized, 
                                        color = estimator_display, group = estimator_display)) +
  geom_line(size = 1.1) +
  geom_point(size = 3) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", size = 0.8) +
  scale_color_manual(values = tte_color_map) +
  labs(
    title = "Standardized Overall RMSE for Subgroup log(AHR) Estimators",
    subtitle = "Standardized to baseline (Subgroup estimator - no shrinkage) RMSE within each scenario",
    x = "Scenario",
    y = "Standardized RMSE (relative to baseline)",
    color = "Estimator"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10),
    legend.position = "bottom",
    legend.text = element_markdown(),
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 2))
```

---

## TTE Table: Mean (Range) of RMSE, Bias, and Coverage by Scenario

```{r tte_table_by_scenario}
# Create summary table with bias/abs_bias ranges and coverage ranges from subgroup metrics
# Filter to main section (exclude Global HN)
tte_table_summary <- tte_subgroup_metrics %>%
  filter(!grepl("Global_HN", estimator)) %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    RMSE_mean = mean(RMSE, na.rm = TRUE),
    RMSE_min = min(RMSE, na.rm = TRUE),
    RMSE_max = max(RMSE, na.rm = TRUE),
    Bias_mean = mean(bias_log, na.rm = TRUE),
    Bias_min = min(bias_log, na.rm = TRUE),
    Bias_max = max(bias_log, na.rm = TRUE),
    AbsBias_mean = mean(abs_bias_log, na.rm = TRUE),
    AbsBias_min = min(abs_bias_log, na.rm = TRUE),
    AbsBias_max = max(abs_bias_log, na.rm = TRUE),
    Coverage_mean = mean(coverage, na.rm = TRUE),
    Coverage_min = min(coverage, na.rm = TRUE),
    Coverage_max = max(coverage, na.rm = TRUE),
    .groups = 'drop'
  )

# Format values and reshape to scenarios as columns, estimators as rows within each metric
tte_table_formatted <- tte_table_summary %>%
  mutate(
    Estimator = beautify_tte_estimator(estimator),
    Scenario = paste("Scenario", scenario_no),
    RMSE_display = paste0(display(RMSE_mean, 2), " (", display(RMSE_min, 2), "-", display(RMSE_max, 2), ")"),
    Bias_display = paste0(display(Bias_mean, 3), " (", display(Bias_min, 3), "-", display(Bias_max, 3), ")"),
    AbsBias_display = paste0(display(AbsBias_mean, 2), " (", display(AbsBias_min, 2), "-", display(AbsBias_max, 2), ")"),
    Coverage_display = paste0(getPercentage(Coverage_mean, 0), " (", getPercentage(Coverage_min, 0), "-", getPercentage(Coverage_max, 0), ")")
  ) %>%
  dplyr::select(Estimator, Scenario, RMSE_display, Bias_display, AbsBias_display, Coverage_display, estimator)

tte_table_formatted <- tte_table_formatted %>%
  mutate(Estimator = ifelse(estimator=="Global_RHS_theta0_delta_plan_half_10_s_2", "Global - RHS(τ₀=δ<sub>plan</sub>/10, s=2)",
                                    Estimator))%>%
  dplyr::select(-estimator)

# Create table grouped by METRIC first, then estimators
tte_pivot_table <- tibble()

# Group 1: RMSE rows
tte_pivot_table <- bind_rows(tte_pivot_table,
  tibble(Metric = "RMSE", Estimator = "", 
         `Scenario 1` = "", `Scenario 2` = "", `Scenario 3` = "", `Scenario 4` = "")
)

for (est in sort(unique(tte_table_formatted$Estimator))) {
  est_data <- tte_table_formatted %>%
    filter(Estimator == est) %>%
    arrange(factor(Scenario, levels = paste("Scenario", 1:4)))
  
  row_data <- tibble(
    Metric = "",
    Estimator = est,
    `Scenario 1` = est_data %>% filter(Scenario == "Scenario 1") %>% pull(RMSE_display) %>% first(),
    `Scenario 2` = est_data %>% filter(Scenario == "Scenario 2") %>% pull(RMSE_display) %>% first(),
    `Scenario 3` = est_data %>% filter(Scenario == "Scenario 3") %>% pull(RMSE_display) %>% first(),
    `Scenario 4` = est_data %>% filter(Scenario == "Scenario 4") %>% pull(RMSE_display) %>% first()
  )
  tte_pivot_table <- bind_rows(tte_pivot_table, row_data)
}

# Group 2: Absolute Bias rows
tte_pivot_table <- bind_rows(tte_pivot_table,
  tibble(Metric = "Absolute Bias", Estimator = "", 
         `Scenario 1` = "", `Scenario 2` = "", `Scenario 3` = "", `Scenario 4` = "")
)


for (est in sort(unique(tte_table_formatted$Estimator))) {
  est_data <- tte_table_formatted %>%
    filter(Estimator == est) %>%
    arrange(factor(Scenario, levels = paste("Scenario", 1:4)))
  
  row_data <- tibble(
    Metric = "",
    Estimator = est,
    `Scenario 1` = est_data %>% filter(Scenario == "Scenario 1") %>% pull(AbsBias_display) %>% first(),
    `Scenario 2` = est_data %>% filter(Scenario == "Scenario 2") %>% pull(AbsBias_display) %>% first(),
    `Scenario 3` = est_data %>% filter(Scenario == "Scenario 3") %>% pull(AbsBias_display) %>% first(),
    `Scenario 4` = est_data %>% filter(Scenario == "Scenario 4") %>% pull(AbsBias_display) %>% first()
  )
  tte_pivot_table <- bind_rows(tte_pivot_table, row_data)
}

# Group 3: Coverage rows
tte_pivot_table <- bind_rows(tte_pivot_table,
  tibble(Metric = "Coverage", Estimator = "", 
         `Scenario 1` = "", `Scenario 2` = "", `Scenario 3` = "", `Scenario 4` = "")
)



for (est in sort(unique(tte_table_formatted$Estimator))) {
  est_data <- tte_table_formatted %>%
    filter(Estimator == est) %>%
    arrange(factor(Scenario, levels = paste("Scenario", 1:4)))
  
  row_data <- tibble(
    Metric = "",
    Estimator = est,
    `Scenario 1` = est_data %>% filter(Scenario == "Scenario 1") %>% pull(Coverage_display) %>% first(),
    `Scenario 2` = est_data %>% filter(Scenario == "Scenario 2") %>% pull(Coverage_display) %>% first(),
    `Scenario 3` = est_data %>% filter(Scenario == "Scenario 3") %>% pull(Coverage_display) %>% first(),
    `Scenario 4` = est_data %>% filter(Scenario == "Scenario 4") %>% pull(Coverage_display) %>% first()
  )
  tte_pivot_table <- bind_rows(tte_pivot_table, row_data)
}

kbl(tte_pivot_table,
    caption = "TTE: Performance metrics (mean and range) by estimator and scenario",
    booktabs = TRUE,
    escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

---

## TTE Table: Worst Subgroup Performance (Scenario 2)

### Performance Table for Subgroup 4a (Worst Subgroup in Scenario 2)

```{r tte_worst_subgroup_performance}
# Subgroup 4a is the worst subgroup in Scenario 2 (AHR = 1.00, highest/least beneficial)
# Create performance table for subgroup 4a by estimator

tte_subgroup_4a_perf <- tte_subgroup_metrics %>%
    filter(!grepl("Global_HN", estimator)) %>%
  filter(scenario_no == "2", join_key == "4a") %>%
  mutate(
    Estimator = beautify_tte_estimator(estimator)
  ) %>%
  dplyr::select(Estimator, RMSE, abs_bias_log, coverage, truth_ahr, n_replications) %>%
  mutate(
    RMSE = display(RMSE, 2),
    AbsBias = display(abs_bias_log, 2),
    Coverage = getPercentage(coverage, 0),
    TruthAHR = display(truth_ahr, 4)
  ) %>%
  dplyr::select(Estimator, RMSE, AbsBias, Coverage) %>%
  arrange(Estimator)

kbl(tte_subgroup_4a_perf,
    caption = "TTE Scenario 2: Performance Metrics for Subgroup 4a (Worst Subgroup, AHR=1.00)",
    booktabs = TRUE,
    escape = FALSE,
    col.names = c("Estimator",  "RMSE", "Absolute Bias", "95% CI Coverage")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

### Worst Subgroup Identification Accuracy

```{r tte_worst_accuracy}
# For each replication and estimator in Scenario 2, identify the subgroup with the highest AHR (worst effect)
# Then check if it correctly identifies subgroup 4a

true_worst_subgroup_4a <- "4a"

predicted_worst <- tte_all_results %>%
    filter(!grepl("Global_HN", estimator)) %>%
  filter(estimator!= "population") %>%
  filter(scenario_no == "2", !is.na(estimate_log)) %>%
  group_by(replication_id, estimator) %>%
  mutate(
    # Convert to HR for comparison (higher AHR = worse treatment effect)
    estimate_ahr = exp(estimate_log),
    is_worst_pred = estimate_ahr == max(estimate_ahr, na.rm = TRUE)
  ) %>%
  filter(is_worst_pred) %>%
  rename(pred_worst_subgroup = join_key) %>%
  distinct()

predicted_worst <- predicted_worst %>%
  mutate(
    error = estimate_log - truth_log,
    ci_contains_truth = (truth_log >= ci_lower) & (truth_log <= ci_upper)
  ) %>%
  group_by(estimator) %>%
  summarise(
    n_replications = n(),
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    bias_log = mean(error, na.rm = TRUE),  # Signed bias per subgroup
    abs_bias_log = abs(bias_log),  # Take absolute value of signed bias (not mean of abs)
    coverage = mean(ci_contains_truth, na.rm = TRUE),
    .groups = 'drop'
  )
# Compare predicted vs true worst subgroups
worst_accuracy <- predicted_worst %>%
  mutate(
    correctly_identified = pred_worst_subgroup == true_worst_subgroup_4a
  ) %>%
  group_by(estimator) %>%
  summarise(
    n_replications = n(),
    n_correct = sum(correctly_identified, na.rm = TRUE),
    accuracy = n_correct / n_replications,
    accuracy_pct = getPercentage(accuracy, 0),
    .groups = 'drop'
  ) %>%
  mutate(
    Estimator = beautify_tte_estimator(estimator)
  ) %>%
  dplyr::select(Estimator, Replications = n_replications, `Correct Identifications` = n_correct, `Accuracy (%)` = accuracy_pct) %>%
  arrange(Estimator)

kbl(worst_accuracy,
    caption = "TTE Scenario 2: Accuracy of Identifying Subgroup 4a as the Worst Subgroup",
    booktabs = TRUE,
    escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE) %>%
  footnote(
    general = "Accuracy shows the proportion of 1,000 replications where each estimator correctly identified subgroup 4a as the worst (highest AHR/least beneficial treatment effect).",
    general_title = ""
  )
```

---

# Continuous Simulation Results for Paper

**NOTE:** This results are the ones excluding X3 from the model fitting (it was excluded from everywhere: data generation and model fitting). Also the Horseshoe model have s=2*sigma_plan
---

## Load Continuous Results

```{r load_continuous_results, include=FALSE}
# Set paths for Continuous
cont_base_path <- "/home/pedreram/bonsaiforest2/simulations/paper/Continuous"
cont_data_file <- file.path(cont_base_path, "continuous_results_prepared.rds")

# Check if prepared data exists
if (!file.exists(cont_data_file)) {
  cat("ERROR: Continuous prepared data file not found!\n")
  cat("Please run: Rscript", file.path(cont_base_path, "prepare_continuous_results.R\n"))
  stop("Missing prepared Continuous data")
}

# Load merged Continuous results (with estimates and truth)
cont_all_results <- readRDS(cont_data_file)

cat("✓ Loaded Continuous merged results from:", cont_data_file, "\n")
cat("  Rows:", nrow(cont_all_results), "\n")
cat("  Estimators:", n_distinct(cont_all_results$estimator), "\n")
cat("  Scenarios:", unique(sort(cont_all_results$scenario_no)), "\n\n")


cont_all_results <- cont_all_results %>%
  dplyr::select(scenario_no, replication_id,subgroup_var, level, estimator, estimate, ci_lower, ci_upper, truth_trt_effect)
```

```{r cont_calculate_metrics}
# Calculate RMSE, Bias from individual replications
# RMSE computed across ALL replications and subgroups combined

# First: calculate per-subgroup metrics for detailed analysis
# IMPORTANT: Compute signed bias first, then take absolute value (like TTE_Results_Analysis.Rmd)
cont_subgroup_metrics <- cont_all_results %>%
  mutate(
    error = estimate - truth_trt_effect
  ) %>%
  group_by(scenario_no, estimator, subgroup_var, level) %>%
  summarise(
    estimate_mean = mean(estimate, na.rm=TRUE),
    estimate_var=var(estimate, na.rm=TRUE),
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    bias = mean(error, na.rm = TRUE),  # Signed bias per subgroup
    abs_bias = abs(bias),  # Take absolute value of signed bias (not mean of abs)
    truth_trt_effect = first(truth_trt_effect),
    .groups = 'drop'
  )

# Second: calculate overall metrics per estimator-scenario
# RMSE is computed from ALL replication-subgroup pairs combined
cont_performance_by_scenario <- cont_subgroup_metrics %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    RMSE = mean(RMSE, na.rm = TRUE),
    Bias = mean(bias, na.rm = TRUE),
    AbsBias = mean(abs_bias, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(as.numeric(scenario_no), estimator)

# Separate datasets: main (exclude Global HN) and supplementary (Global HN only)
cont_is_global_hn <- grepl("Global_HN", cont_performance_by_scenario$estimator)
cont_performance_main <- cont_performance_by_scenario[!cont_is_global_hn, ]
cont_performance_supplementary <- cont_performance_by_scenario[cont_is_global_hn, ]

# Overall metrics across all scenarios
cont_performance_overall <- cont_all_results %>%
  mutate(
    error = estimate - truth_trt_effect
  ) %>%
  group_by(estimator, scenario_no) %>%
  summarise(
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    Bias = mean(error, na.rm = TRUE),
    AbsBias = mean(abs(error), na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(estimator)

cat("Continuous performance metrics calculated.\n")
```

---

## Continuous Helper Functions

```{r cont_helpers}
# Beautify estimator names for Continuous
beautify_cont_estimator <- function(estimator_name) {
  dplyr::case_when(
    estimator_name == "population_with_covariates" ~ "Population with covariates",
    estimator_name == "population" ~ "Population",
    estimator_name == "subgroup" ~ "Subgroup",
    # Global HN
    grepl("Global_HN.*phi_delta_half", estimator_name) ~ "Global - HN(φ=δ<sub>plan</sub>/2)",
    grepl("Global_HN.*phi_sigma_plan", estimator_name) ~ "Global - HN(φ=σ<sub>plan</sub>)",
    grepl("Global_HN.*phi_delta_plan$", estimator_name) ~ "Global - HN(φ=δ<sub>plan</sub>)",
    # Global Horseshoe
        grepl("Global_Horseshoe.*delta_10", estimator_name) ~ "Global - Horseshoe(τ₀ =δ<sub>plan</sub>/10)",
            grepl("Global_Horseshoe.*delta_100", estimator_name) ~ "Global - Horseshoe(τ₀ =δ<sub>plan</sub>/100)",
    grepl("Global_Horseshoe.*delta_half", estimator_name) ~ "Global - Horseshoe(τ₀ =δ<sub>plan</sub>/2)",
    grepl("Global_Horseshoe.*sigma_plan", estimator_name) ~ "Global - Horseshoe(τ₀ =σ<sub>plan</sub>)",
    grepl("Global_Horseshoe.*delta_plan$", estimator_name) ~ "Global - Horseshoe(τ₀ =δ<sub>plan</sub>)",
            # OVAT HN Cova Adjusted
    grepl("OVAT_HN.*phi_delta_halfCov", estimator_name) ~ "OVATCOV - HN(φ=δ<sub>plan</sub>/2)",
    grepl("OVAT_HN.*phi_sigma_planCov$", estimator_name) ~ "OVATCOV - HN(φ=σ<sub>plan</sub>)",
    grepl("OVAT_HN.*phi_delta_planCov$", estimator_name) ~ "OVATCOV - HN(φ=δ<sub>plan</sub>)",
    # OVAT HN
    grepl("OVAT_HN.*phi_delta_half", estimator_name) ~ "OVAT - HN(φ=δ<sub>plan</sub>/2)",
    grepl("OVAT_HN.*phi_sigma_plan", estimator_name) ~ "OVAT - HN(φ=σ<sub>plan</sub>)",
    grepl("OVAT_HN.*phi_delta_plan$", estimator_name) ~ "OVAT - HN(φ=δ<sub>plan</sub>)",
    TRUE ~ estimator_name
  )
}
```

---

## Continuous Figure: Standardized Overall RMSE

```{r cont_figure_rmse, fig.width=12, fig.height=6}
# Prepare data: standardize RMSE within each scenario by baseline (subgroup estimator)
# Filter to main section (exclude Global HN)
cont_plot_data <- cont_performance_main %>%
  mutate(
    scenario_no_numeric = as.numeric(scenario_no),
    estimator_display = beautify_cont_estimator(estimator)
  )

# Use subgroup estimator as baseline for standardization
cont_baselines <- cont_performance_by_scenario %>%
  filter(estimator == "subgroup") %>%
  dplyr::select(scenario_no, baseline_rmse = RMSE)


# Standardize RMSE
cont_plot_data_standardized <- cont_plot_data %>%
  left_join(cont_baselines, by = "scenario_no") %>%
  mutate(
    RMSE_standardized = RMSE/baseline_rmse 
  )

# Create plot with manual color scheme
cont_color_map <- c(
  "Subgroup" = "red",
  "Population" = "orange",
  "Population with covariates" = "yellow",
  "OVAT - HN(φ=δ<sub>plan</sub>/2)" = "skyblue",
  "OVAT - HN(φ=δ<sub>plan</sub>)" = "royalblue",
  "OVAT - HN(φ=σ<sub>plan</sub>)" = "navy",
  "OVATCOV - HN(φ=δ<sub>plan</sub>/2)" = "violet",
  "OVATCOV - HN(φ=δ<sub>plan</sub>)" = "purple",
  "OVATCOV - HN(φ=σ<sub>plan</sub>)" = "purple4",
  "Global - Horseshoe(τ₀ =δ<sub>plan</sub>/10)" = "palegreen3",
  "Global - Horseshoe(τ₀ =δ<sub>plan</sub>)" = "forestgreen",
  "Global - Horseshoe(τ₀ =σ<sub>plan</sub>)" = "darkgreen"
)

ggplot(cont_plot_data_standardized, aes(x = scenario_no_numeric, y = RMSE_standardized, 
                                        color = estimator_display, group = estimator_display)) +
  geom_line(size = 1.1) +
  geom_point(size = 3) +
  #geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", size = 0.8) +
  scale_x_continuous(breaks = 1:3, limits = c(0.8, 3.2)) +
  scale_color_manual(values = cont_color_map) +
  labs(
    title = "Standardized Overall RMSE for Subgroup Treatment Effect Estimators",
    subtitle = "Standardized to baseline (Subgroup estimator - no shrinkage) RMSE within each scenario",
    x = "Scenario",
    y = "Standardized RMSE (relative to baseline)",
    color = "Estimator"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10),
    legend.position = "bottom",
    legend.text = element_markdown(),
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 4))
```

---

## Continuous Table: Mean (Range) of RMSE, Bias, and Coverage by Scenario

```{r cont_table_by_scenario}
# Create summary table with bias/abs_bias ranges from subgroup metrics
# Filter to main section (exclude Global HN)

cont_table_summary <- cont_subgroup_metrics %>%
  filter(!grepl("Global_HN", estimator)) %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    RMSE_mean = mean(RMSE, na.rm = TRUE),
    RMSE_min = min(RMSE, na.rm = TRUE),
    RMSE_max = max(RMSE, na.rm = TRUE),
    Bias_mean = mean(bias, na.rm = TRUE),
    Bias_min = min(bias, na.rm = TRUE),
    Bias_max = max(bias, na.rm = TRUE),
    AbsBias_mean = mean(abs_bias, na.rm = TRUE),
    AbsBias_min = min(abs_bias, na.rm = TRUE),
    AbsBias_max = max(abs_bias, na.rm = TRUE),
    .groups = 'drop'
  )

# First: calculate per-subgroup metrics with coverage if available
cont_subgroup_metrics_full <- cont_all_results %>%
  mutate(
    error = estimate - truth_trt_effect,
    ci_contains_truth = if_else(!is.na(ci_lower) & !is.na(ci_upper),
                                 (truth_trt_effect >= ci_lower) & (truth_trt_effect <= ci_upper),
                                 NA)
  ) %>%
  group_by(scenario_no, estimator, subgroup_var, level) %>%
  summarise(
    n_replications = n(),
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    abs_bias = abs(mean(error, na.rm = TRUE)),
    coverage = if_else(all(is.na(ci_contains_truth)), NA_real_, mean(ci_contains_truth, na.rm = TRUE)),
    truth_trt_effect = first(truth_trt_effect),
    .groups = 'drop'
  )

# Update cont_table_summary to include coverage and remove bias
cont_table_summary <- cont_subgroup_metrics_full %>%
  filter(!grepl("Global_HN", estimator)) %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    RMSE_mean = mean(RMSE, na.rm = TRUE),
    RMSE_min = min(RMSE, na.rm = TRUE),
    RMSE_max = max(RMSE, na.rm = TRUE),
    AbsBias_mean = mean(abs_bias, na.rm = TRUE),
    AbsBias_min = min(abs_bias, na.rm = TRUE),
    AbsBias_max = max(abs_bias, na.rm = TRUE),
    Coverage_mean = mean(coverage, na.rm = TRUE),
    Coverage_min = min(coverage, na.rm = TRUE),
    Coverage_max = max(coverage, na.rm = TRUE),
    .groups = 'drop'
  )

# Format values and reshape to scenarios as columns, estimators as rows within each metric
cont_table_formatted <- cont_table_summary %>%
  mutate(
    Estimator = beautify_cont_estimator(estimator),
    Scenario = paste("Scenario", scenario_no),
    RMSE_display = paste0(display(RMSE_mean, 2), " (", display(RMSE_min, 2), "-", display(RMSE_max, 2), ")"),
    AbsBias_display = paste0(display(AbsBias_mean, 2), " (", display(AbsBias_min, 2), "-", display(AbsBias_max, 2), ")"),
    Coverage_display = paste0(getPercentage(Coverage_mean, 0), " (", getPercentage(Coverage_min, 0), "-", getPercentage(Coverage_max, 0), ")")
  ) %>%
  dplyr::select(Estimator,estimator ,Scenario, RMSE_display, AbsBias_display, Coverage_display)

cont_table_formatted <- cont_table_formatted %>%
  mutate(Estimator = ifelse(estimator=="Global_Horseshoe_global_delta_100", "Global - Horseshoe(τ_0 =δ_plan/100, s=2*σ_plan)",
                                    Estimator))%>%
  dplyr::select(-estimator)

# Create table grouped by METRIC first, then estimators
cont_pivot_table <- tibble()

# Group 1: RMSE rows
cont_pivot_table <- bind_rows(cont_pivot_table,
  tibble(Metric = "RMSE", Estimator = "", 
         `Scenario 1` = "", `Scenario 2` = "", `Scenario 3` = "")
)

for (est in sort(unique(cont_table_formatted$Estimator))) {
  est_data <- cont_table_formatted %>%
    filter(Estimator == est) %>%
    arrange(factor(Scenario, levels = paste("Scenario", 1:3)))
  
  row_data <- tibble(
    Metric = "",
    Estimator = est,
    `Scenario 1` = est_data %>% filter(Scenario == "Scenario 1") %>% pull(RMSE_display) %>% first(),
    `Scenario 2` = est_data %>% filter(Scenario == "Scenario 2") %>% pull(RMSE_display) %>% first(),
    `Scenario 3` = est_data %>% filter(Scenario == "Scenario 3") %>% pull(RMSE_display) %>% first()
  )
  cont_pivot_table <- bind_rows(cont_pivot_table, row_data)
}

# Group 2: Absolute Bias rows
cont_pivot_table <- bind_rows(cont_pivot_table,
  tibble(Metric = "Absolute Bias", Estimator = "", 
         `Scenario 1` = "", `Scenario 2` = "", `Scenario 3` = "")
)

for (est in sort(unique(cont_table_formatted$Estimator))) {
  est_data <- cont_table_formatted %>%
    filter(Estimator == est) %>%
    arrange(factor(Scenario, levels = paste("Scenario", 1:3)))
  
  row_data <- tibble(
    Metric = "",
    Estimator = est,
    `Scenario 1` = est_data %>% filter(Scenario == "Scenario 1") %>% pull(AbsBias_display) %>% first(),
    `Scenario 2` = est_data %>% filter(Scenario == "Scenario 2") %>% pull(AbsBias_display) %>% first(),
    `Scenario 3` = est_data %>% filter(Scenario == "Scenario 3") %>% pull(AbsBias_display) %>% first()
  )
  cont_pivot_table <- bind_rows(cont_pivot_table, row_data)
}

# Group 3: Coverage rows
cont_pivot_table <- bind_rows(cont_pivot_table,
  tibble(Metric = "Coverage", Estimator = "", 
         `Scenario 1` = "", `Scenario 2` = "", `Scenario 3` = "")
)

for (est in sort(unique(cont_table_formatted$Estimator))) {
  est_data <- cont_table_formatted %>%
    filter(Estimator == est) %>%
    arrange(factor(Scenario, levels = paste("Scenario", 1:3)))
  
  row_data <- tibble(
    Metric = "",
    Estimator = est,
    `Scenario 1` = est_data %>% filter(Scenario == "Scenario 1") %>% pull(Coverage_display) %>% first(),
    `Scenario 2` = est_data %>% filter(Scenario == "Scenario 2") %>% pull(Coverage_display) %>% first(),
    `Scenario 3` = est_data %>% filter(Scenario == "Scenario 3") %>% pull(Coverage_display) %>% first()
  )
  cont_pivot_table <- bind_rows(cont_pivot_table, row_data)
}

kbl(cont_pivot_table,
    caption = "Continuous: Performance metrics (mean and range) by estimator and scenario",
    booktabs = TRUE,
    escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

---

## Continuous Forest Plots: RMSE by Subgroup and Scenario

```{r cont_forest_plot_prep}
# Prepare data for forest plots: use per-subgroup RMSE
# Use main estimators (exclude Global HN) for cleaner visualization

cont_forest_data <- cont_subgroup_metrics %>%
  filter(!grepl("Global_HN", estimator)) %>%
  mutate(
    estimator_display = beautify_cont_estimator(estimator),
    subgroup_label = paste0(subgroup_var, "=", level)
  ) %>%
  arrange(as.numeric(scenario_no), estimator_display, subgroup_label)

cat("Forest plot data prepared:\n")
cat("  Scenarios:", unique(cont_forest_data$scenario_no), "\n")
cat("  Estimators:", n_distinct(cont_forest_data$estimator_display), "\n")
cat("  Subgroups per scenario:", n_distinct(cont_forest_data$subgroup_label), "\n")
```

```{r cont_forest_plot}
# Create forest plots for each scenario
# Traditional forest plot style showing RMSE per subgroup

scenarios <- sort(unique(cont_forest_data$scenario_no))

# Use the same color map as the overall RMSE plot
cont_color_map <- c(
  "Subgroup" = "red",
  "Population" = "orange",
  "OVAT - HN(φ=δ<sub>plan</sub>/2)" = "skyblue",
  "OVAT - HN(φ=δ<sub>plan</sub>)" = "royalblue",
  "OVAT - HN(φ=σ<sub>plan</sub>)" = "navy",
  "Global - Horseshoe(τ₀ =δ<sub>plan</sub>/10, s=2*σ<sub>plan</sub>)" = "palegreen1",
  "Global - Horseshoe(τ₀ =δ<sub>plan</sub>/2, s=2*σ<sub>plan</sub>)" = "palegreen3",
  "Global - Horseshoe(τ₀ =σ<sub>plan</sub>, s=2*σ<sub>plan</sub>)" = "forestgreen",
  "Global - Horseshoe(τ₀ =δ<sub>plan</sub>, s=2*σ<sub>plan</sub>)" = "darkgreen"
)

for (scen in scenarios) {
  scen_data <- cont_forest_data %>%
    filter(scenario_no == scen) %>%
    # Create a combined label for sorting/grouping
    mutate(
      subgroup_label_only = paste0(subgroup_var, "=", level)
    ) %>%
    # Sort by subgroup for better grouping
    arrange(subgroup_var, level, estimator_display) %>%
    mutate(
      subgroup_label_only = factor(subgroup_label_only, 
                                    levels = unique(subgroup_label_only))
    )
  
  # Create forest plot
  p <- ggplot(scen_data, aes(x = RMSE, y = subgroup_label_only, color = estimator_display)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "gray70", size = 0.5) +
    geom_point(size = 4, alpha = 0.8, position = position_dodge(width = 0.5)) +
    scale_color_manual(values = cont_color_map, name = "Estimator") +
    labs(
      title = paste0("Continuous Scenario ", scen, ": RMSE per Subgroup"),
      subtitle = "Smaller RMSE indicates better estimation accuracy",
      x = "Root Mean Squared Error (RMSE)",
      y = "Subgroup"
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(face = "bold", size = 13),
      plot.subtitle = element_text(size = 10, color = "gray50"),
      axis.text.y = element_text(size = 10),
      axis.text.x = element_text(size = 10),
      legend.position = "right",
      legend.text = element_markdown(size = 9),
      panel.grid.major.y = element_line(color = "gray95", size = 0.3),
      panel.grid.minor.x = element_blank(),
      panel.spacing = unit(1, "lines")
    )
  
  print(p)
}
```

---

## Continuous Table: Performance for Scenario 3 (Strong Heterogeneity)

### Performance Table for X11cat level a (Worst Subgroup in Scenario 3)

```{r cont_scenario_3_analysis}
# Scenario 3 typically has strong heterogeneity
# X11cat level a is the worst subgroup (largest absolute effect)
# Create performance table for X11cat=a by estimator using cont_subgroup_metrics_full

cont_subgroup_x11cat_a_perf <- cont_subgroup_metrics_full %>%
  filter(!grepl("Global_HN", estimator)) %>%
  filter(scenario_no == "3", subgroup_var == "X11cat", level == "a") %>%
  mutate(
    Estimator = beautify_cont_estimator(estimator)
  ) %>%
  dplyr::select(Estimator, RMSE, abs_bias, coverage, truth_trt_effect, n_replications) %>%
  mutate(
    RMSE = display(RMSE, 2),
    AbsBias = display(abs_bias, 2),
    Coverage = getPercentage(coverage, 0),
    TruthEffect = display(truth_trt_effect, 4)
  ) %>%
  dplyr::select(Estimator, RMSE, AbsBias, Coverage) %>%
  arrange(Estimator)

kbl(cont_subgroup_x11cat_a_perf,
    caption = "Continuous Scenario 3: Performance Metrics for X11cat level a (Worst Subgroup)",
    booktabs = TRUE,
    escape = FALSE,
    col.names = c("Estimator", "RMSE", "Absolute Bias", "95% CI Coverage")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

### Worst Subgroup Identification Accuracy

```{r cont_worst_accuracy}
# For each replication and estimator in Scenario 3, identify the subgroup with the smallest absolute treatment effect (closest to 0 = worst)
# Then check if it correctly identifies X11cat=a

true_worst_subgroup_a <- "a"
true_worst_subgroup_var <- "X11cat"

predicted_worst_cont <- cont_all_results %>%
  filter(!grepl("Global_HN", estimator)) %>%
  filter(estimator != "population") %>%
  filter(scenario_no == "3", !is.na(estimate)) %>%
  group_by(replication_id, estimator) %>%
  mutate(
    # Identify worst: 
    is_worst_pred = estimate == min(estimate, na.rm = TRUE)
  ) %>%
  filter(is_worst_pred) %>%
  dplyr::select(scenario_no, replication_id, estimator, subgroup_var, level) %>%
  rename(pred_worst_var = subgroup_var, pred_worst_level = level) %>%
  distinct()

# Compare predicted vs true worst subgroups (should be X11cat=a)
worst_accuracy_cont <- predicted_worst_cont %>%
  mutate(
    correctly_identified = (pred_worst_var == true_worst_subgroup_var) & (pred_worst_level == true_worst_subgroup_a)
  ) %>%
  group_by(estimator) %>%
  summarise(
    n_replications = n(),
    n_correct = sum(correctly_identified, na.rm = TRUE),
    accuracy = n_correct / n_replications,
    accuracy_pct = getPercentage(accuracy, 0),
    .groups = 'drop'
  ) %>%
  mutate(
    Estimator = beautify_cont_estimator(estimator)
  ) %>%
  dplyr::select(Estimator, Replications = n_replications, `Correct Identifications` = n_correct, `Accuracy (%)` = accuracy_pct) %>%
  arrange(Estimator)

kbl(worst_accuracy_cont,
    caption = "Continuous Scenario 3: Accuracy of Identifying X11cat level a as the Worst Subgroup",
    booktabs = TRUE,
    escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE) %>%
  footnote(
    general = "Accuracy shows the proportion of 1,000 replications where each estimator correctly identified X11cat=a as the worst (smallest absolute treatment effect - closest to 0).",
    general_title = ""
  )
```

---

# Supplementary: Global HN Model Results

This section presents detailed results for the Global Horseshoe-Normal (HN) models with varying prior specifications.

---

## TTE: Global HN Model Performance

### TTE Figure: Global HN Standardized RMSE

```{r tte_supp_figure_global_hn, fig.width=10, fig.height=6, eval= FALSE}
# Prepare supplementary data: only Global HN models
tte_plot_data_supp <- tte_performance_supplementary %>%
  mutate(
    scenario_no_numeric = as.numeric(scenario_no),
    estimator_display = beautify_tte_estimator(estimator),
    scenario_label = beautify_tte_scenario(scenario_no)
  )

# Use subgroup estimator as baseline for standardization
tte_baselines_supp <- tte_performance_by_scenario %>%
  filter(estimator == "subgroup") %>%
  dplyr::select(scenario_no, baseline_rmse = RMSE)

# Standardize RMSE
tte_plot_data_supp_standardized <- tte_plot_data_supp %>%
  left_join(tte_baselines_supp, by = "scenario_no") %>%
  mutate(
    RMSE_standardized = RMSE / baseline_rmse
  )

# Create plot with manual color scheme
tte_supp_color_map <- c(
  "Global - HN(φ=1)" = "lightgray",
  "Global - HN(φ=δ<sub>plan</sub>/2)" = "gray50",
  "Global - HN(φ=δ<sub>plan</sub>)" = "darkgray"
)

ggplot(tte_plot_data_supp_standardized, aes(x = scenario_no_numeric, y = RMSE_standardized, 
                                             color = estimator_display, group = estimator_display)) +
  geom_line(size = 1.1) +
  geom_point(size = 3) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", size = 0.8) +
  scale_x_continuous(breaks = 1:4, limits = c(0.8, 4.2)) +
  scale_color_manual(values = tte_supp_color_map) +
  labs(
    title = "Standardized RMSE for Global Horseshoe-Normal HN Models",
    subtitle = "TTE: subgroup log(AHR) estimation across scenarios",
    x = "Scenario",
    y = "Standardized RMSE (relative to subgroup baseline)",
    color = "HN Prior"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10),
    legend.position = "bottom",
    legend.text = element_markdown(),
    panel.grid.minor = element_blank()
  )
```

### TTE Table: Global HN Performance by Scenario

```{r tte_supp_table_global_hn, eval= FALSE}
# Create summary table for Global HN only
tte_table_supp_summary <- tte_subgroup_metrics %>%
  filter(grepl("Global_HN", estimator)) %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    RMSE_mean = mean(RMSE, na.rm = TRUE),
    RMSE_min = min(RMSE, na.rm = TRUE),
    RMSE_max = max(RMSE, na.rm = TRUE),
    Bias_mean = mean(bias_log, na.rm = TRUE),
    Bias_min = min(bias_log, na.rm = TRUE),
    Bias_max = max(bias_log, na.rm = TRUE),
    AbsBias_mean = mean(abs_bias_log, na.rm = TRUE),
    AbsBias_min = min(abs_bias_log, na.rm = TRUE),
    AbsBias_max = max(abs_bias_log, na.rm = TRUE),
    Coverage_mean = mean(coverage, na.rm = TRUE),
    .groups = 'drop'
  )

tte_supp_table_data <- tte_table_supp_summary %>%
  mutate(
    Estimator = beautify_tte_estimator(estimator),
    RMSE_display = paste0(display(RMSE_mean, 2), " (", display(RMSE_min, 2), "-", display(RMSE_max, 2), ")"),
    Bias_range = paste0(display(Bias_mean, 3), " (", display(Bias_min, 3), "-", display(Bias_max, 3), ")"),
    AbsBias_range = paste0(display(AbsBias_mean, 2), " (", display(AbsBias_min, 2), "-", display(AbsBias_max, 2), ")"),
    Coverage = getPercentage(Coverage_mean, 0)
  ) %>%
  dplyr::select(
    Scenario = scenario_no,
    Estimator,
    RMSE = RMSE_display,
    Bias = Bias_range,
    `|Bias|` = AbsBias_range,
    Coverage
  ) %>%
  arrange(as.numeric(Scenario), Estimator)

kbl(tte_supp_table_data,
    caption = "TTE Supplementary: Global HN models - Overall RMSE and mean (range) of bias across all replications and subgroups",
    booktabs = TRUE,
    escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE) %>%
  collapse_rows(columns = 1)
```

---

## Continuous: Global HN Model Performance

### Continuous Figure: Global HN Standardized RMSE

```{r cont_supp_figure_global_hn, fig.width=10, fig.height=6, eval= FALSE}
# Prepare supplementary data: only Global HN models
cont_plot_data_supp <- cont_performance_supplementary %>%
  mutate(
    scenario_no_numeric = as.numeric(scenario_no),
    estimator_display = beautify_cont_estimator(estimator)
  )

# Use subgroup estimator as baseline for standardization
cont_baselines_supp <- cont_performance_by_scenario %>%
  filter(estimator == "subgroup") %>%
  dplyr::select(scenario_no, baseline_rmse = RMSE)

# Standardize RMSE
cont_plot_data_supp_standardized <- cont_plot_data_supp %>%
  left_join(cont_baselines_supp, by = "scenario_no") %>%
  mutate(
    RMSE_standardized = RMSE / baseline_rmse
  )

# Create plot with manual color scheme
cont_supp_color_map <- c(
  "Global - HN(φ=δ<sub>plan</sub>/2)" = "lightgray",
  "Global - HN(φ=δ<sub>plan</sub>)" = "gray50",
  "Global - HN(φ=σ<sub>plan</sub>)" = "darkgray"
)

ggplot(cont_plot_data_supp_standardized, aes(x = scenario_no_numeric, y = RMSE_standardized, 
                                              color = estimator_display, group = estimator_display)) +
  geom_line(size = 1.1) +
  geom_point(size = 3) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", size = 0.8) +
  scale_x_continuous(breaks = 1:3, limits = c(0.8, 3.2)) +
  scale_color_manual(values = cont_supp_color_map) +
  labs(
    title = "Standardized RMSE for Global Horseshoe-Normal HN Models",
    subtitle = "Continuous: subgroup treatment effect estimation across scenarios",
    x = "Scenario",
    y = "Standardized RMSE (relative to subgroup baseline)",
    color = "HN Prior"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 11),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 10),
    legend.position = "bottom",
    legend.text = element_markdown(),
    panel.grid.minor = element_blank()
  )
```

### Continuous Table: Global HN Performance by Scenario

```{r cont_supp_table_global_hn, eval= FALSE}
# Create summary table for Global HN only
cont_table_supp_summary <- cont_subgroup_metrics %>%
  filter(grepl("Global_HN", estimator)) %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    RMSE_mean = mean(RMSE, na.rm = TRUE),
    RMSE_min = min(RMSE, na.rm = TRUE),
    RMSE_max = max(RMSE, na.rm = TRUE),
    Bias_mean = mean(bias, na.rm = TRUE),
    Bias_min = min(bias, na.rm = TRUE),
    Bias_max = max(bias, na.rm = TRUE),
    AbsBias_mean = mean(abs_bias, na.rm = TRUE),
    AbsBias_min = min(abs_bias, na.rm = TRUE),
    AbsBias_max = max(abs_bias, na.rm = TRUE),
    .groups = 'drop'
  )

cont_supp_table_data <- cont_table_supp_summary %>%
  mutate(
    Estimator = beautify_cont_estimator(estimator),
    RMSE_display = paste0(display(RMSE_mean, 2), " (", display(RMSE_min, 2), "-", display(RMSE_max, 2), ")"),
    Bias_range = paste0(display(Bias_mean, 4), " (", display(Bias_min, 4), "-", display(Bias_max, 4), ")"),
    AbsBias_range = paste0(display(AbsBias_mean, 2), " (", display(AbsBias_min, 2), "-", display(AbsBias_max, 2), ")")
  ) %>%
  dplyr::select(
    Scenario = scenario_no,
    Estimator,
    RMSE = RMSE_display,
    Bias = Bias_range,
    `|Bias|` = AbsBias_range
  ) %>%
  arrange(as.numeric(Scenario), Estimator)

kbl(cont_supp_table_data,
    caption = "Continuous Supplementary: Global HN models - Overall RMSE and mean (range) of bias across all replications and subgroups",
    booktabs = TRUE,
    escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE) %>%
  collapse_rows(columns = 1)
```

---

**Document generated**: `r Sys.time()`
````

**Document generated**: `r Sys.time()`
