---
title: "Advanced Functionalities"
author: Miriam Pedrera Gomez, Isaac Gravestock, and Marcel Wolbers
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    number_sections: true
    citation_package: natbib
    base_format: rmarkdown::html_vignette
bibliography: "references.bib"
link-citations: true   
linkcolor: blue
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{Advanced Functionalities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

This vignette demonstrates advanced functionalities supported by `bonsaiforest2` and the underlying `brms` package, including adding an **offset to negative binomial regression models** for counts, **customizing prior distributions**, and **applying stratification** to nuisance parameters.

# Handling Exposure in Count Outcomes with Offsets

For count data, the observed value often depends on the **exposure time** or **follow-up time**. Using an **offset** variable is the standard statistical method to properly account for this time variation by modeling the event **rate** (count per unit time) instead of the raw count.

In `bonsaiforest2`, you can include the offset directly in the `response_formula` using the syntax: `count ~ trt + offset(log_fup_duration)`.

### Example 1: Count outcome with offset {-}

The example uses the `shrink_data` dataset to illustrate a negative binomial model for variable `count` accounting for treatment `trt` and the subgrouping variables `x1`, `x2`, and `x3`. Each subject's follow-up time is provided in variable `fup_duration` whose log-transformed value is included as an offset. 

```{r}
# Load packages and data
library(bonsaiforest2)
library(brms)

shrink_data <- bonsaiforest2::shrink_data

# prepare offset variable
shrink_data$log_fup_duration <- log(shrink_data$fup_duration)
```

```{r, fig.width=8, fig.height=10}
count_model_fit <- run_brms_analysis(
  data = shrink_data,
  response_type = "count",
  response_formula = count ~ trt + offset(log_fup_duration),
  unshrunk_terms_formula  = ~ 1 + x1 + x2 + x3,
  shrunk_predictive_formula = ~ 0 + trt:x1 + trt:x2 + trt:x3,
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

count_model_summary <- summary_subgroup_effects(brms_fit = count_model_fit)

plot(count_model_summary, title = "Global model for count data: All subgrouping variables")
```

# Customizing Prior Distributions

In Bayesian analysis, the choice of priors is important and can materially affect the results.
For most settings, the `brms` default of flat priors for the intercept (`intercept_prior`) and unshrunk terms (`unshrunk_prior`, includes the prior for the main treatment effect) are sensible and do not need fine-tuning. However, **we recommend that the user carefully considers the prior for shrunken terms**, in particular those for covariate-by-treatment interactions (**`shrunk_predictive_prior`**). 

`bonsaiforest2` gives the user full flexibility to specify priors as described via examples below. 

The **default priors** used by `bonsaiforest2` are as follows: 

| Prior Component | Parameter Name | Default Prior | 
|:---|:---|:---|:---|
| **Intercept** | `intercept_prior` | `NULL` (i.e. `brms` default) | 
| **Unshrunk Terms** | `unshrunk_prior` | `NULL` (i.e. `brms` default) | 
| **Shrunk Prognostic** | `shrunk_prognostic_prior` | `horseshoe(scale_global = 1)`; in case the formula specifies random effects (pipe-pipe syntax), automatically uses a normal prior with a half-normal hyperprior with $\phi=1$ for the standard deviation instead|
| **Shrunk Predictive** | `shrunk_predictive_prior` | `horseshoe(scale_global = 1)`; in case the formula specifies random effects (pipe-pipe syntax), automatically uses a normal prior with a half-normal hyperprior with $\phi=1$ for the standard deviation instead|

The following examples demonstrate how to customize priors in `bonsaiforest2`. We use the `shrink_data` package dataset with a time-to-event outcome to show different prior specification strategies, from simple defaults to advanced hierarchical structures.

### Model Preparation

```{r}
# Prepare basic model formulas
prepared_model <- prepare_formula_model(
  data = shrink_data,
  response_type = "survival",
  response_formula = Surv(tt_event, event_yn) ~ trt,
  unshrunk_terms_formula = ~ x1 + x2,
  shrunk_predictive_formula = ~ 0 + trt:x1
)
```

### Example 2: Using priors directly supported by `brms` {-}

The simplest approach: use the priors directly supported by `brms` to adjust priors of different terms. 
Below, we assume that the trial targeted a hazard ratio of 0.7, i.e. $\delta_{plan}=|log(0.7)$, and set `scale_global` of the regularized horseshoe prior to this value.   

```{r}
fit_ex2 <- fit_brms_model(
  prepared_model = prepared_model,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_predictive_prior = "horseshoe(scale_global = abs(log(0.7)), scale_slab = 2, df_slab = 4)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

# View the priors that were automatically set
cat("\n=== Priors Used ===\n")
print(fit_ex2[["prior"]])
```

### Example 3: Using the R2-D2 shrinkage prior {-}

An alternative global-local shrinkage prior to the regularized horseshoe prior is the R2-D2 prior. 
In the linear regression setting, it can be motivated as specifying a prior on the model's coefficient of determination ($R^2$) first, and then distributing the prior through to the coefficients. 
The R2-D2 prior is also directly supported by `brms` as illustrated below. 

```{r}
fit_ex3 <- fit_brms_model(
  prepared_model = prepared_model,
  shrunk_predictive_prior = "R2D2(mean_R2 = 0.5, prec_R2 = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

cat("\n=== Priors Used ===\n")
print(fit_ex3[["prior"]])
```

### Example 4: Custom hierarchical prior (advanced) {-}

This example demonstrates injecting raw Stan code using `stanvars`. This is necessary if you want to implement a hierarchical structure that `brms` does not support natively, such as estimating a shared variance parameter across coefficients.

```{r}
# 1. Define new hyperparameters in Stan
stanvars_full_hierarchical <- brms::stanvar(
  scode = "  real mu_pred;\n  real<lower=0> sigma_pred;\n",
  block = "parameters"
) +
  # Add priors for these parameters
  brms::stanvar(
    scode = "  // Priors on the hierarchical parameters\n  target += normal_lpdf(mu_pred | 0, 4); \n  target += normal_lpdf(sigma_pred | 0, 1) - normal_lccdf(0 | 0, 1); \n",
    block = "model"
  )

# 2. Create prior that references the Stan variables
prior_full_hierarchical <- brms::set_prior("normal(mu_pred, sigma_pred)")

# 3. Pass both to fit_brms_model
fit_ex4 <- fit_brms_model(
  prepared_model = prepared_model,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = prior_full_hierarchical,
  stanvars = stanvars_full_hierarchical,
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

# View the used priors
cat("\n=== Priors Used ===\n")
print(fit_ex4[["prior"]])
print(fit_ex4[["stanvars"]])
```

### Example 5: Coefficient-specific priors {-}

By default, `bonsaiforest2` uses the same prior for all coefficients assigned to the group of "unshrunken terms", "shrunken prognostic terms", or "shrunken predictive terms", respectively. However, it is possible to fine-tune this by assigning a different prior to one or several coefficients within a group. 

The example below illustrates this. It includes the interaction `trt*x3` as an "unshrunken term" and assigns a tighter prior to the regression coefficients for `trt:x3b`, `trt:x3c`, and `trt:x3d`, respectively, compared to other coefficients in the same group. 

To implement this, we first call `prepare_formula_model()` to extract the exact coefficient names that will be included in the model.
Then, we define coefficient-specific priors for these by combining `brms::set_prior` calls.  


```{r}
# 1. Run prepare_formula_model
prepared_model_ex5 <- prepare_formula_model(
  data = shrink_data,
  response_type = "survival",
  response_formula = Surv(tt_event, event_yn) ~ trt,
  unshrunk_terms_formula = ~ x1 + trt*x3,
  shrunk_predictive_formula = ~ 0 + trt:x1,
)

# 2. Inspect the Results
# A. The generated brms formula object
print(prepared_model_ex5$formula)

# B. The processed terms
print(prepared_model_ex5$stan_variable_names$X_unshrunktermeffect)

cat("=== Prior Strategy ===\n")
cat("General unshrunk prior: normal(0, 5)\n")
cat("Specific for trt:x3 interactions: normal(0, 1)\n\n")

# Create combined prior object
# IMPORTANT: Use EXACT coefficient names from the prepared data
unshrunk_priors_combined <- c(
  brms::set_prior("normal(0, 5)", class = "b"),  # General
  brms::set_prior("normal(0, 1)", class = "b", coef = "trt:x3b"),
  brms::set_prior("normal(0, 1)", class = "b", coef = "trt:x3c"),
  brms::set_prior("normal(0, 1)", class = "b", coef = "trt:x3d")
)

# Fit the model
fit_ex5 <- fit_brms_model(
  prepared_model = prepared_model_ex5,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = unshrunk_priors_combined,  # Pass the combined object
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

# View the used priors
cat("\n=== Priors Used ===\n")
print(fit_ex5[["prior"]])

```

### Example 6: Hierarchical priors with shared variance (stanvars) {-}

Create correlated priors by sharing a common variance parameter estimated from the data. This is sensible when you believe treatment effects across subgroups are exchangeable and want to borrow information across them. 

```{r}
cat("\n=== Hierarchical Prior Structure ===\n")
cat("Instead of independent priors per coefficient, we pool information through a shared scale:\n")
cat("  tau ~ half-normal(0, 1)  [shared scale parameter]\n")
cat("  beta_i ~ N(0, tau) for each coefficient i\n")
cat("This creates exchangeability: we assume coefficients are similar but with adaptive variation.\n\n")

# Step 1: Declare tau as a parameter to be estimated
tau_parameter <- brms::stanvar(
  scode = "  real<lower=0> biomarker_tau;  // Shared scale parameter\n",
  block = "parameters"
)

# Step 2: Add prior for tau (using normal truncated to positive values with constraint)
tau_prior <- brms::stanvar(
  scode = "  biomarker_tau ~ normal(0, 1);  // Hyperprior on the scale\n",
  block = "model"
)

# Combine stanvars
hierarchical_stanvars <- tau_parameter + tau_prior

# Step 3: Create priors referencing the shared variance parameter
# Note: We identified these coefficient names using prepare_formula_model above
unshrunk_priors_hier <- c(
  brms::set_prior("normal(0, 5)", class = "b"),  # General
  brms::set_prior("normal(0, biomarker_tau)", class = "b", coef = "trt:x3b"),
  brms::set_prior("normal(0, biomarker_tau)", class = "b", coef = "trt:x3c"),
  brms::set_prior("normal(0, biomarker_tau)", class = "b", coef = "trt:x3d")
)
```

```{r, eval=FALSE}
# Step 4: Fit the hierarchical model
fit_ex6 <- fit_brms_model(
  prepared_model = prepared_model_ex5,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = unshrunk_priors_hier,
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  stanvars = hierarchical_stanvars,
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

# View the used priors
cat("\n=== Priors Used ===\n")
print(fit_ex6[["prior"]])
print(fit_ex6[["stanvars"]])
```

<!-- # Custom contrasts -->

<!-- !! Note: The "custom contrasts" example below appears to apply Helmert contrasts to the unshrunk term x2 but -->
<!-- !!       for trt:x2, one hot encoding is still used; hence it is commented for now                           -->

<!-- By default, `bonsaiforest2` uses one hot encoding shrunken predictive terms specified with `~ 0 + trt:x1 + ...`, i.e., we include an indicator variable for each level of the categorical covariate-by-treatment interaction in the design matrix accepting an overparameterized model. The rationale for this is that we want to treat all levels of the categorical covariates symmetrically. In contrast, The design matrix corresponding to unshrunken terms uses dummy encoding (`contr.treatment`), i.e., we omit baseline levels of categorical variables and absorb them in the intercept. -->

<!-- However, it is possible to manually set custom contrasts for your data using `contrasts()`, and the library will preserve your choice throughout the analysis. -->


<!-- ### Example 7: Custom contrast encoding for shrunken terms {-} -->


<!-- *Scenario*: You want to use custom contrast coding for subgroup variables, such as Helmert contrasts or sum contrasts, instead of the default treatment contrasts. This is useful for specific hypothesis structures or when comparing non-orthogonal effects. -->

<!-- *Key Technique*: Set contrasts using `contrasts()` before calling `prepare_formula_model()`. The library will preserve your choice. -->

<!-- ```{r} -->
<!-- # Use the shrink_data dataset -->
<!-- sample_data_contrast <- bonsaiforest2::shrink_data -->

<!-- # Set Helmert contrasts for the x2 variable (3 levels: a, b, c) -->
<!-- contrasts(sample_data_contrast$x2) <- contr.helmert(3) -->
<!-- cat("Contrast matrix for x2:\n") -->
<!-- print(contrasts(sample_data_contrast$x2)) -->

<!-- # Prepare model -->
<!-- prepared_custom_contrast <- prepare_formula_model( -->
<!--   data = sample_data_contrast, -->
<!--   response_formula = y ~ trt, -->
<!--   unshrunk_terms_formula = ~ x2, -->
<!--   shrunk_predictive_formula = ~ 0 + trt:x2, -->
<!--   response_type = "continuous" -->
<!-- ) -->

<!-- # Observe custom contrast in the data -->
<!-- str(prepared_custom_contrast$data) -->

<!-- # Fit the model -->
<!-- fit_custom_contrast <- fit_brms_model( -->
<!--   prepared_model = prepared_custom_contrast, -->
<!--   intercept_prior = "normal(0, 5)", -->
<!--   unshrunk_prior = "normal(0, 2.5)", -->
<!--   shrunk_predictive_prior = "horseshoe(scale_global = 1)", -->
<!--   chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr" -->
<!-- ) -->

<!-- estimate_custom_contrast <- summary_subgroup_effects(fit_custom_contrast) -->

<!-- print(estimate_custom_contrast) -->
<!-- ``` -->

<!-- # Prediction on new data -->

<!-- !! Note: Currently, brms does not supports actual prediction of survival times, so it's not quite clear what -->
<!-- !!       exactly is predicted here and why this is useful; hence it is commented for now                     -->


<!-- After fitting a model, you can use `brms::predict()` to generate predictions of the linear predictor for new subjects- -->


<!-- ### Example 7: Prediction on new data {-} -->

<!-- We use the model fitted in example 6 to predict outcomes for new patients with different characteristics. -->

<!-- ```{r} -->
<!-- # Use the fitted model from Example 7 -->
<!-- # Generate new data for prediction -->
<!-- set.seed(789) -->
<!-- new_patients <- data.frame( -->
<!--   id = 1:20, -->
<!--   trt = rep(0:1, length.out = 20), -->
<!--   x2 = factor(rep(c("a", "b", "c"), length.out = 20), levels = c("a", "b", "c")) -->
<!-- ) -->

<!-- # IMPORTANT: Set the same contrasts as in Example 7 -->
<!-- # The prediction data must use the same encoding as the training data -->
<!-- contrasts(new_patients$x2) <- contr.helmert(3) -->

<!-- # Prepare new data -->
<!-- prepared_custom_contrast <- prepare_formula_model( -->
<!--   data = new_patients, -->
<!--   response_formula = y ~ trt, -->
<!--   shrunk_predictive_formula = ~ 0 + trt:x2,  -->
<!--   response_type = "continuous" -->
<!-- ) -->

<!-- # Generate predictions using brms::predict -->
<!-- # This returns posterior predictive samples -->
<!-- predictions <- predict(fit_custom_contrast, newdata = prepared_custom_contrast$data, summary = TRUE) -->

<!-- head(predictions) -->
<!-- ``` -->

# Stratification for nuisance parameters

Stratified models which assume different residual variances (continuous endpoints), overdispersion parameters (count endpoints), or baseline hazards (time-to-event endpoints) across strata are also supported in `bonsaiforest2` (via `brms`). The argument `stratification_formula` defines the grouping factor(s). 

### Example 7: Fully stratified continuous one-way shrinkage model {-}

We fit a one-way model for the continuous outcome `y` in `shrink_data` with subgroups defined by the binary covariate `x1` with adjustment for `x2` and `x3` as prognostic factors.

In the model specification below, we assume separate regression coefficients for `x2` and `x3`, respectively, across strata defined by `x1` (specified via interaction terms `x1*x2` and `x1*x3`) as well as different residual variances across strata (specified via argument `stratification_formula`). The only parameter that is shared across strata is the shrinkage parameter for treatment effect heterogeneity. 

```{r}
# model with x1*x2 and x1*x3 (separate regression coef per level of x1 for x2 and x3), and separate variance per x1 strata
oneway_x1_flex_strat <- run_brms_analysis(
  data = shrink_data,
  response_type = "continuous",
  response_formula = y ~ trt,
  stratification_formula = ~ x1,
  unshrunk_terms_formula = ~ x1 + x1*x2 + x1*x3, 
  shrunk_predictive_formula = ~ (0 + trt || x1),
  shrunk_predictive_prior = set_prior("normal(0, 0.3)", class = "sd"),
  chains = 2, iter = 2000, warmup = 1000, cores = 2,
  refresh = 0, backend = "cmdstanr"
)
```

```{r}
summary_oneway_x1_flex_strat <- summary_subgroup_effects(
  brms_fit = oneway_x1_flex_strat
)

print(summary_oneway_x1_flex_strat$estimates)
```

### Example 8: Stratified one-way survival model {-}

Assume that the primary analysis of the trial was stratified by variable `x2`. 
A corresponding one-way model for `x1` which uses a Cox model stratified by `x2` is fitted below. 


```{r, fig.width=9, fig.height=4}
# Model fitting with stratified baseline hazard
fit_surv_oneway_x1_strat <- run_brms_analysis(
  data = shrink_data,
  response_type = "survival",
  response_formula = Surv(tt_event, event_yn) ~ trt,
  stratification_formula = ~ x2,
  unshrunk_terms_formula = ~ x1 + x3, 
  shrunk_predictive_formula = ~ (0 + trt || x1),
  shrunk_predictive_prior = set_prior("normal(0, abs(log(0.7)))", class = "sd"),
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

summary_surv_oneway_x1_strat <- summary_subgroup_effects(fit_surv_oneway_x1_strat)

print(summary_surv_oneway_x1_strat$estimates)

plot(summary_surv_oneway_x1_strat)
```

# Summary

This vignette demonstrated several advanced functionalities supported by `bonsaiforest2`:

1.  **Offset variables** for count outcomes with varying exposure times
2.  **Custom prior specification** with flexible prior constraints
3.  **Stratification** for nuisance parameters that vary across groups

These features provide researchers with powerful tools for complex trial analyses while maintaining the principled Bayesian shrinkage framework.
