---
title: "Advanced Functionalities"
author: Miriam Pedrera Gomez, Isaac Gravestock, and Marcel Wolbers
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    number_sections: true
    citation_package: natbib
    base_format: rmarkdown::html_vignette
bibliography: "references.bib"
link-citations: true   
linkcolor: blue
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{Advanced Functionalities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

This vignette demonstrates advanced functionalities for Bayesian shrinkage estimation, including specifying an **offset for count data**, customizing **prior distributions**, and using **stratification** to handle heterogeneity in nuisance parameters.

# Handling Exposure in Count Outcomes with Offsets

For count data (like disease exacerbations or event rates), the observed count often depends on the **exposure time** or **follow-up time**. Using an **offset** variable is the standard statistical method to properly account for this time variation by modeling the event **rate** (count per unit time) instead of the raw count.

In `bonsaiforest2`, you can include the offset directly in the `response_formula` using the standard `brms` syntax: `y + offset(log_exposure_time) ~ trt`.

## Example 1: Count Outcome with Offset (Disease Exacerbations)

This scenario models exacerbation counts using a Negative Binomial distribution and explicitly accounts for the patient's exposure time.

*Scenario*: Modeling count outcomes with exposure-time adjustment. Adjust for `baseline` (unshrunk prognostic effect) and multiple exploratory `biomarkers` (shrunk prognostic effects). Overall treatment effect on the rate scale.

```{r}
# Data Simulation 
set.seed(789)
library(bonsaiforest2)
set.seed(789)
n <- 200

# Create biomarker data with simple naming
biomarker_data <- as.data.frame(matrix(rnorm(n * 8), ncol = 8))
names(biomarker_data) <- paste0("B", 1:8)

# Generate count data with treatment effect heterogeneity
count_data <- data.frame(
  y = rnbinom(n, size = 1.5, mu = 3),
  trt = rep(0:1, length.out = n),
  baseline = rnorm(n, 10, 2),
  log_exposure_time = log(runif(n, 0.5, 1.5))
)

# Add biomarkers and account for heterogeneous treatment effects
count_data <- cbind(count_data, biomarker_data)
count_data$y <- rnbinom(n, size = 1.5, mu = exp(0.5 + 0.1 * count_data$baseline + 
                        count_data$trt * (0.3 + 0.2 * (count_data$B1 > 0))))

# Create formula string for all biomarkers
shrunk_prog_str <- paste("~ 0 +", paste(names(biomarker_data), collapse = " + "))
```

```{r}
# Model Fitting with Offset
count_model_fit <- run_brms_analysis(
  data = count_data,
  # Include offset(log_exposure_time) directly in the response formula
  response_formula = y + offset(log_exposure_time) ~ trt,
  response_type = "count",
  unshrunk_terms_formula = ~ baseline,
  shrunk_prognostic_formula = as.formula(shrunk_prog_str),
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)
```

# Customizing Prior Distributions

The choice of prior is central to Bayesian shrinkage. `bonsaiforest2` provides sensible defaults, but it allows for full customization using dedicated prior arguments: `intercept_prior`, `unshrunk_prior`, `shrunk_prognostic_prior`, and `shrunk_predictive_prior`.

## Prior Specification Mechanics

Priors are specified using separate parameters for each component.

| Prior Component | Parameter Name | Default Prior | Notes |
|:---|:---|:---|:---|
| **Intercept** | `intercept_prior` | `NULL` (brms default) | Optional custom prior |
| **Unshrunk Terms** | `unshrunk_prior` | `NULL` (brms default) | Optional custom prior |
| **Shrunk Prognostic** | `shrunk_prognostic_prior` | `horseshoe(scale_global = 1)` | Strong shrinkage for fixed effects (colon syntax); For random effects (pipe-pipe syntax), automatically uses `normal(0, 1)` on SD scale |
| **Shrunk Predictive** | `shrunk_predictive_prior` | `horseshoe(scale_global = 1)` | Strong shrinkage for fixed effects (colon syntax); For random effects (pipe-pipe syntax), can be specified via `set_prior("normal(0, 1)", class = "sd")` |

## Practical Examples of Prior Setting

The following examples demonstrate how to customize priors using the new API. We use a synthetic survival dataset to show different prior specification strategies, from simple defaults to advanced hierarchical structures.

### Dataset Generation and Model Preparation

This dataset mimics a clinical trial with treatment effects that vary by subgroups.

```{r}
# Load library
library(bonsaiforest2)

# 1. Create Sample Data with treatment heterogeneity
set.seed(123)
n <- 200

sim_data <- data.frame(
  time = round(runif(n, 1, 100)),
  status = sample(0:1, n, replace = TRUE),
  trt = rep(0:1, length.out = n),
  age = rnorm(n, 50, 10),
  X1 = factor(sample(c("A", "B"), n, replace = TRUE)),
  X2 = factor(sample(c("A", "B", "C"), n, replace = TRUE))
)

# Ensure variables are factors
sim_data$trt <- factor(sim_data$trt, levels = c(0, 1))
sim_data$X1 <- as.factor(sim_data$X1)
sim_data$X2 <- as.factor(sim_data$X2)

# 2. Prepare the model formula
prepared_model <- prepare_formula_model(
  data = sim_data,
  response_formula = Surv(time, status) ~ trt,
  shrunk_predictive_formula = ~ 0 + trt:X1,
  unshrunk_terms_formula = ~ age,
  response_type = "survival"
)
```

### Example 2: Using Default Priors

The simplest approach: use the package defaults that adapt to your outcome type (horseshoe(1) for shrunk effects, brms defaults for others).

```{r}
fit_ex3 <- fit_brms_model(
  prepared_model = prepared_model,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

# View the priors that were automatically set
cat("\n=== Priors Used ===\n")
print(fit_ex3[["prior"]])
```

### Example 3: Using the R2D2 Shrinkage Prior

The R2D2 prior is useful when you want to control the *global* shrinkage via the coefficient of determination ($R^2$) rather than a scale parameter. This is often more interpretable for stakeholders.

```{r}
# Use a custom R2D2 prior for the shrunk predictive effects (interactions)

fit_ex4 <- fit_brms_model(
  prepared_model = prepared_model,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = "R2D2(mean_R2 = 0.5, prec_R2 = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

cat("\n=== Priors Used ===\n")
print(fit_ex4[["prior"]])
```

### Example 4: Custom Hierarchical Prior (Advanced)

This example demonstrates injecting raw Stan code using `stanvars`. This is necessary if you want to implement a hierarchical structure that `brms` does not support natively, such as estimating a shared variance parameter across coefficients.

```{r}
# 1. Define new hyperparameters in Stan
stanvars_full_hierarchical <- brms::stanvar(
  scode = "  real mu_pred;\n  real<lower=0> sigma_pred;\n",
  block = "parameters"
) +
  # Add priors for these parameters
  brms::stanvar(
    scode = "  // Priors on the hierarchical parameters\n  target += normal_lpdf(mu_pred | 0, 4); \n  target += normal_lpdf(sigma_pred | 0, 1) - normal_lccdf(0 | 0, 1); \n",
    block = "model"
  )

# 2. Create prior that references the Stan variables
prior_full_hierarchical <- brms::set_prior("normal(mu_pred, sigma_pred)")

# 3. Pass both to fit_brms_model
fit_ex5 <- fit_brms_model(
  prepared_model = prepared_model,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = prior_full_hierarchical,
  stanvars = stanvars_full_hierarchical,
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

# View the used priors
cat("\n=== Priors Used ===\n")
print(fit_ex5[["prior"]])
print(fit_ex5[["stanvars"]])
```

### Example 5: Coefficient-Specific Priors

You can set different priors for specific coefficients by passing a `brmsprior` object (created with `c()`).

*Scenario*: Set a general prior for all unshrunk terms, but use a tighter prior specifically for treatment-biomarker interactions.

*Key Technique*: Use `prepare_formula_model()` to discover the exact coefficient names that will appear in the model.

```{r}
# 1. Run prepare_formula_model
prepared_model <- prepare_formula_model(
  data = sim_data,
  response_formula = Surv(time, status) ~ trt,
  shrunk_predictive_formula = ~ 0 + trt:X1,
  unshrunk_terms_formula = ~ age + trt*X2,
  response_type = "survival"
)

# 2. Inspect the Results
# A. The generated brms formula object
print(prepared_model$formula)

# B. The processed terms
print(prepared_model$stan_variable_names$X_unshrunktermeffect)

cat("=== Prior Strategy ===\n")
cat("General unshrunk prior: normal(0, 5)\n")
cat("Specific for trt:X2 interactions: normal(0, 1)\n\n")

# Create combined prior object
# IMPORTANT: Use EXACT coefficient names from the prepared data
# These names appear after contrast encoding in prepare_formula_model
unshrunk_priors_combined <- c(
  brms::set_prior("normal(0, 5)", class = "b"),  # General
  brms::set_prior("normal(0, 1)", class = "b", coef = "trt:X2B"),
  brms::set_prior("normal(0, 1)", class = "b", coef = "trt:X2C")
)

# Fit the model
fit_specific <- fit_brms_model(
  prepared_model = prepared_model,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = unshrunk_priors_combined,  # Pass the combined object
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

# View the used priors
cat("\n=== Priors Used ===\n")
print(fit_specific[["prior"]])

```

### Example 6: Hierarchical Priors with Shared Variance (stanvars)

Create correlated priors by sharing a common variance parameter estimated from the data.

*Use case*: When you believe treatment effects across subgroups should be exchangeable, you can pool information by giving them a shared variance.

```{r}
cat("\n=== Hierarchical Prior Structure ===\n")
cat("Instead of independent priors per coefficient, we pool information through a shared scale:\n")
cat("  tau ~ half-normal(0, 1)  [shared scale parameter]\n")
cat("  beta_i ~ N(0, tau) for each coefficient i\n")
cat("This creates exchangeability: coefficients are ~similar but with adaptive variation.\n\n")

# Step 1: Declare tau as a parameter to be estimated
tau_parameter <- brms::stanvar(
  scode = "  real<lower=0> biomarker_tau;  // Shared scale parameter\n",
  block = "parameters"
)

# Step 2: Add prior for tau (using normal truncated to positive values with constraint)
tau_prior <- brms::stanvar(
  scode = "  biomarker_tau ~ normal(0, 1);  // Hyperprior on the scale\n",
  block = "model"
)

# Combine stanvars
hierarchical_stanvars <- tau_parameter + tau_prior

# Step 3: Create priors referencing the shared variance parameter
# Note: We identified these coefficient names using prepare_formula_model above
unshrunk_priors_hier <- c(
  brms::set_prior("normal(0, 5)", class = "b"),  # General
  brms::set_prior("normal(0, biomarker_tau)", class = "b", coef = "trt:X2B"),
  brms::set_prior("normal(0, biomarker_tau)", class = "b", coef = "trt:X2C")
)
```

```{r, eval=FALSE}
# Step 4: Fit the hierarchical model
fit_hier <- fit_brms_model(
  prepared_model = prepared_model,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = unshrunk_priors_hier,
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  stanvars = hierarchical_stanvars,
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

# View the used priors
cat("\n=== Priors Used ===\n")
print(fit_hier[["prior"]])
print(fit_hier[["stanvars"]])
```

# Advanced Model Parameterization

This section demonstrates advanced features for controlling how your model represents subgroups.

## Example 7: Custom Contrast Encoding for Shrunk Terms

By default, `bonsaiforest2` uses one-hot encoding (all factor levels, no reference category) for shrunk interaction terms specified with `~ 0 + ...` syntax, enabling full hierarchical shrinkage of all levels. For unshrunk terms, treatment contrasts (k-1 dummy variables with a reference level) are used by default. However, you can manually set custom contrasts in your data using `contrasts()`, and the library will preserve your choice throughout the analysis.

*Scenario*: You want to use custom contrast coding for subgroup variables, such as Helmert contrasts or sum contrasts, instead of the default treatment contrasts. This is useful for specific hypothesis structures or when comparing non-orthogonal effects.

*Key Technique*: Set contrasts using `contrasts()` before calling `prepare_formula_model()`. The library will preserve your choice.

```{r}
# Create sample data with multiple subgroups
set.seed(456)
n <- 250
sample_data_contrast <- data.frame(
  id = 1:n,
  trt = rep(0:1, length.out = n),
  y = rnorm(n, 50, 10),
  baseline = rnorm(n, 65, 10),
  X1 = factor(sample(c("A", "B", "C"), n, replace = TRUE))
)

# Set Helmert contrasts for the X1 variable
contrasts(sample_data_contrast$X1) <- contr.helmert(3)
cat("Contrast matrix for X1:\n")
print(contrasts(sample_data_contrast$X1))

# Prepare model 
prepared_custom_contrast <- prepare_formula_model(
  data = sample_data_contrast,
  response_formula = y ~ trt,
  unshrunk_terms_formula = ~ baseline,
  shrunk_predictive_formula = ~ 0 + trt:X1, 
  response_type = "continuous"
)

# Observe costum contrast in the data
str(prepared_custom_contrast$data)

# Fit the model
fit_custom_contrast <- fit_brms_model(
  prepared_model = prepared_custom_contrast,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)

estimate_custom_contrast <- summary_subgroup_effects(fit_custom_contrast)

print(estimate_custom_contrast)
```

## Example 8: Prediction on New Data

After fitting a model, you can use `brms::predict()` to generate predictions for new patients.

*Scenario*: We've fitted a treatment effect model and want to predict outcomes for new patients with different characteristics.

```{r}
# Use the fitted model from Example 7
# Generate new data for prediction
set.seed(789)
new_patients <- data.frame(
  id = 1:20,
  trt = rep(0:1, length.out = 20),
  baseline = rnorm(20, 65, 10),
  X1 = factor(rep(c("A", "B", "C"), length.out = 20))
)

# IMPORTANT: Set the same contrasts as in Example 7
# The prediction data must use the same encoding as the training data
contrasts(new_patients$X1) <- contr.helmert(3)

# Prepare new data
prepared_custom_contrast <- prepare_formula_model(
  data = new_patients,
  response_formula = y ~ trt,
  unshrunk_terms_formula = ~ baseline,
  shrunk_predictive_formula = ~ 0 + trt:X1, 
  response_type = "continuous"
)

# Generate predictions using brms::predict
# This returns posterior predictive samples
predictions <- predict(fit_custom_contrast, newdata = prepared_custom_contrast$data, summary = TRUE)

head(predictions)
```

# Stratification for Nuisance Parameters

In many trials, parameters like the observation error variance ($\sigma^2$ for continuous outcomes) or the baseline hazard function ($h_0(t)$ for survival outcomes) are known to vary by site, country, or other factors. Stratification models this known heterogeneity by fitting these nuisance parameters separately for each level of a grouping variable.

Use the `stratification_formula` argument to define the grouping factor(s).

## Example 9: Stratified Continuous Model

*Scenario*: We model outcomes where the observation error **standard deviation** $\sigma$ differs by site. Stratification uses `brms` distributional formulas to estimate separate residual variance parameters for each site, allowing for heterogeneity in measurement noise or outcome variability across sites.

```{r}
set.seed(42)
n <- 250
sigma_by_site <- c(A = 6, B = 12, C = 18)
sample_data_strat_cont <- data.frame(
    id = 1:n,
    site = factor(sample(c("A", "B", "C"), n, replace = TRUE))
)
sample_data_strat_cont$trt <- rep(0:1, length.out = n)
sample_data_strat_cont$baseline <- rnorm(n, mean = 100, sd = 8)
sample_data_strat_cont$X1 <- factor(sample(c("A", "B"), n, replace = TRUE))

# Generate outcome with treatment heterogeneity and site-specific noise
noise <- rnorm(n, mean = 0, sd = sigma_by_site[sample_data_strat_cont$site])
sample_data_strat_cont$y <- 50 + 
                            sample_data_strat_cont$trt * (-8 + 2 * (as.numeric(sample_data_strat_cont$X1) - 1)) +
                            0.2 * (sample_data_strat_cont$baseline - 100) +
                            noise
```

```{r}
# Model Fitting with Stratified Model

fit_continuous_stratified <- run_brms_analysis(
  data = sample_data_strat_cont,
  response_formula = y ~ trt,
  response_type = "continuous",
  unshrunk_terms_formula = ~ baseline,
  shrunk_predictive_formula = ~ 0 + trt:X1,
  stratification_formula = ~ site,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)
```

```{r}
strat_continuous_summary <- summary_subgroup_effects(
  brms_fit = fit_continuous_stratified
  # All parameters automatically extracted!
)

print(strat_continuous_summary$estimates)
plot(strat_continuous_summary, title = "Stratified Continuous: Subgroup Effects")
```

## Example 10: Stratified Survival Model

*Scenario*: We model a time-to-event outcome with stratified baseline hazard by country using a piecewise exponential model. Stratification via `stratification_formula = ~ country` allows the baseline hazard to differ by country, accommodating regional differences in baseline risk or standard of care.

```{r}
set.seed(123)
n <- 250
lambda_by_country <- c(A = 0.01, B = 0.03)
surv_data_strat <- data.frame(
    id = 1:n,
    country = factor(sample(c("A", "B"), n, replace = TRUE)),
    trt = rep(0:1, length.out = n),
    age = rnorm(n, 65, 10),
    X1 = factor(sample(c("A", "B"), n, replace = TRUE))
)

# Linear predictor with treatment heterogeneity
lp <- (as.numeric(surv_data_strat$trt) - 1) * -0.6 + 
      (surv_data_strat$age - 65) * 0.03 +
      (as.numeric(surv_data_strat$trt) - 1) * (as.numeric(surv_data_strat$X1) - 1) * -0.5

# Generate event times
u <- runif(n)
lambda_vec <- lambda_by_country[surv_data_strat$country]
gamma <- 1.5 # Weibull shape parameter
true_event_time <- (-log(u) / (lambda_vec * exp(lp)))^(1/gamma)
censoring_time <- 60 # Administrative censoring time
surv_data_strat$event_status <- ifelse(true_event_time <= censoring_time, 1, 0)
surv_data_strat$event_time <- pmin(true_event_time, censoring_time)
```

```{r}
# Model Fitting with Stratified Baseline Hazard
fit_surv_stratified <- run_brms_analysis(
  data = surv_data_strat,
  response_formula = Surv(event_time, event_status) ~ trt,
  response_type = "survival",
  unshrunk_terms_formula = ~ age,
  shrunk_predictive_formula = ~ 0 + trt:X1,
  stratification_formula = ~ country,
  intercept_prior = "normal(0, 5)",
  unshrunk_prior = "normal(0, 2.5)",
  shrunk_prognostic_prior = "horseshoe(scale_global = 1)",
  shrunk_predictive_prior = "horseshoe(scale_global = 1)",
  chains = 2, iter = 1000, warmup = 500, cores = 2, refresh = 0, backend = "cmdstanr"
)
```

```{r ex6-summary-plot}
strat_surv_summary <- summary_subgroup_effects(
  brms_fit = fit_surv_stratified
  # All parameters automatically extracted!
)

print(strat_surv_summary$estimates)
plot(strat_surv_summary, title = "Stratified Survival: Subgroup Effects (AHR)")
```

# Summary

This vignette demonstrated advanced functionalities in `bonsaiforest2`:

1.  **Offset variables** for count outcomes with varying exposure times
2.  **Custom prior specification** with flexible prior constraints
3.  **One-hot encoding** for full factor representation in interactions
4.  **Coefficient-specific priors** for fine-grained control
5.  **Hierarchical priors** with shared variance using `stanvars`
6.  **Stratification** for nuisance parameters that vary by groups

These features provide researchers with powerful tools for complex trial analyses while maintaining the principled Bayesian shrinkage framework.
