---
title: "Quickstart"
author: Miriam Pedrera Gomez, Isaac Gravestock, and Marcel Wolbers
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    number_sections: true
    citation_package: natbib
    base_format: rmarkdown::html_vignette
bibliography: "references.bib"
link-citations: true   
linkcolor: blue
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Introduction

The `bonsaiforest2` package consists of 3 core functions which are typically called in sequence:

1.  `run_brms_analysis()` - Prepares the model formula and fits the Bayesian model using `brms`.
2.  `summary_subgroup_effects()` - Calculates the marginal subgroup treatment effects.
3.  `plot()` - Creates a forest plot from the summary object.

The package enables the implementation of both **global models** [@wolbers2025using], that estimate all prognostic and predictive effects in a single unified model, and **one-way models** [@wang2024bayesian], that estimate treatment fitting a different model for each subgrouping variable.

This vignette demonstrates how to use the package to fit and compare these different modeling formulas. You'll learn how to:

-   Fit one-way models (random effects with treatment slopes varying by subgroup)
-   Fit global models (all subgroup variables in one unified model)
-   Generate summaries of subgroup treatment effects
-   Visualize and compare results from different model specifications

This example makes use of Bayesian modeling, which requires the installation of the [`brms`](https://paulbuerkner.com/brms/) package and a working Stan installation (e.g., via [`cmdstanr`](https://mc-stan.org/cmdstanr/)).

# The Data

We will use the `shrink_data` package dataset with a continuous response variable. The relevant outcome is `y`.

We consider a model where we want to find the treatment effect (`trt`) on `y`. We explore **three subgroup variables** as **predictive** variables (potential treatment effect modifiers):

-   `x1`: Subgroup 1 (categories: a, b)
-   `x2`: Subgroup 2 (categories: a, b, c)
-   `x3`: Subgroup 3 (categories: a, b, c, d)

First, let's load the libraries and the data.

```{r load-packages, message=FALSE, warning=FALSE}
# Load the main package
library(bonsaiforest2)
library(brms)

```

```{r ex-data}
# Load the shrink_data package dataset
shrink_data <- bonsaiforest2::shrink_data

print(head(shrink_data))
```

# One-way Models

This section demonstrates how to fit separate models, each examining one subgroup variable at a time. In one-way models, we specify treatment effects as random slopes using pipe-pipe notation (e.g., `~ (0 + trt || subgroup)`), which allows treatment effects to vary by subgroup with automatic hierarchical regularization via random effects. Note that for each model we also include always the subgrouping variable we want to test the treatment effects for also as prognostic.

## One-way Model: x1 Only

```{r oneway-x1, message=FALSE, warning=FALSE}
# Fit model with only x1 as subgroup variable using one-way approach
# Random effects notation (0 + trt || x1) estimates varying treatment slopes by x1
oneway_x1 <- run_brms_analysis(
  data = shrink_data,
  response_type = "continuous",
  response_formula = y ~ trt,
  unshrunk_terms_formula = ~ x1,
  shrunk_prognostic_formula = NULL,
  shrunk_predictive_formula = ~ (0 + trt || x1),
  intercept_prior = "normal(0, 10)",
  shrunk_predictive_prior = set_prior("normal(0, 1)", class = "sd"),
  chains = 2, iter = 2000, warmup = 1000, cores = 2,
  refresh = 0, backend = "cmdstanr"
)

summary_oneway_x1 <- summary_subgroup_effects(brms_fit = oneway_x1)
print(summary_oneway_x1)
```

## One-way Model: x2 Only

```{r oneway-x2, message=FALSE, warning=FALSE}
oneway_x2 <- run_brms_analysis(
  data = shrink_data,
  response_type = "continuous",
  response_formula = y ~ trt,
  shrunk_prognostic_formula = NULL,
    unshrunk_terms_formula = ~ x2,
  shrunk_predictive_formula = ~ (0 + trt || x2),
  intercept_prior = "normal(0, 10)",
  shrunk_predictive_prior = set_prior("normal(0, 1)", class = "sd"),
  chains = 2, iter = 2000, warmup = 1000, cores = 2,
  refresh = 0, backend = "cmdstanr"
)

summary_oneway_x2 <- summary_subgroup_effects(brms_fit = oneway_x2)
```

## One-way Model: x3 Only

```{r oneway-x3, message=FALSE, warning=FALSE}
oneway_x3 <- run_brms_analysis(
  data = shrink_data,
  response_type = "continuous",
  response_formula = y ~ trt,
  unshrunk_terms_formula = ~ x3,
  shrunk_prognostic_formula = NULL,
  shrunk_predictive_formula = ~ (0 + trt || x3),
  intercept_prior = "normal(0, 10)",
  shrunk_predictive_prior = set_prior("normal(0, 1)", class = "sd"),
  chains = 2, iter = 2000, warmup = 1000, cores = 2,
  refresh = 0, backend = "cmdstanr"
)

summary_oneway_x3 <- summary_subgroup_effects(brms_fit = oneway_x3)
```

## One-way Models: Visualizing All Models

You can combine and visualize results from multiple models using `combine_summaries()`:

```{r compare-all-oneway, fig.width=9, fig.height=8}
# Combine all one-way models
combined_oneway <- combine_summaries(list(
  "x1" = summary_oneway_x1,
  "x2" = summary_oneway_x2,
  "x3" = summary_oneway_x3
))

plot(combined_oneway, title = "One-way Models: All Subgroup Variables")
```

# Global Model

This section demonstrates how to fit a single model that includes all subgroup variables simultaneously using the global approach. All treatment-by-subgroup interactions are estimated in one unified model with colon notation (e.g., `~ 0 + trt:subgroup`) and strong regularization (Horseshoe prior) applied to the interaction terms.

## Global Model: All Subgroups

```{r global-fit, message=FALSE, warning=FALSE}
# Fit a single unified model with ALL subgroup variables simultaneously using global approach
# - Unshrunk prognostic effects: subgroup main effects without shrinkage
# - Shrunk predictive effects: treatment interactions with strong regularization using one-hot encoding
global_shrinkage_model <- run_brms_analysis(
  data = shrink_data,
  response_type = "continuous",
  response_formula = y ~ trt,
  unshrunk_terms_formula  = ~ 0 + x1 + x2 + x3,
  shrunk_predictive_formula = ~ 0 + trt:x1 + trt:x2 + trt:x3,
  intercept_prior = "normal(0, 10)",
  shrunk_predictive_prior = "horseshoe(scale_global = 0.5)",
  chains = 2, iter = 2000, warmup = 1000, cores = 2,
  refresh = 0, backend = "cmdstanr"
)
```

## Global Model: Summary of Subgroup Effects

Use `summary_subgroup_effects()` to generate marginal treatment effects for each subgroup. The function automatically extracts all treatment interactions from the fitted model:

```{r global-summary}
global_summary <- summary_subgroup_effects(brms_fit = global_shrinkage_model)

# Print the summary of subgroup-specific treatment effects
print(global_summary)
```

## Global Model: Visualization

Use the `plot()` function to create a forest plot from the summary object:

```{r global-plot, fig.width=8, fig.height=10}
plot(global_summary, title = "Global Model: All Subgroup Variables")
```

# Comparing Multiple Models in One Plot

The `plot()` function supports comparing multiple models side-by-side. Pass a named list of `subgroup_summary` objects to create a comparative forest plot.

## Example: Comparing One-way vs Global Models

```{r compare-models, fig.width=9, fig.height=7}
# Combine summaries for comparison
combined <- combine_summaries(list(
  "One-way" = combined_oneway,
  "Global" = global_summary
))

# Plot the comparison
plot(combined, title = "Comparing One-way vs Global Models")
```
