---
title: "Continuous Simulation Results Analysis"
subtitle: "RMSE, Bias, and Covariate Balance Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                      fig.width = 14, fig.height = 9)

library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(rlang)
```

# Continuous Simulation Results Analysis

This document analyzes Continuous outcome simulation results, calculating:
- **RMSE** (Root Mean Squared Error)
- **Bias** (Mean difference between estimate and truth)
- **Coverage** (Proportion of 95% CIs containing the true value)

All metrics are calculated on the original scale (treatment effect).

---

## Load Results

```{r load_results}
# Set paths
base_path <- "/home/pedreram/bonsaiforest2/simulations/Continuous"
results_path <- file.path(base_path, "Results")

# Find all RDS files
all_files <- list.files(results_path, pattern = "\\.rds$", full.names = TRUE)
cat("Found", length(all_files), "result files:\n")
cat(paste("-", basename(all_files), collapse = "\n"), "\n")

```

---

## Load Truth Values

```{r load_truth}
# Load truth data
truth_file <- file.path(base_path, "Scenarios", "truth.RData")
load(truth_file)
cat("Loaded truth values from:", truth_file, "\n")

# Extract population-level truth (treatment effect)
truth_population <- simulation_truth %>%
  group_by(scenario_no, scenario) %>%
  summarise(
    truth_trt_effect = first(pop_trt_effect),
    .groups = 'drop'
  ) %>%
  mutate(scenario_no = as.character(scenario_no))

cat("\nTruth data loaded successfully.\n")
cat("Scenarios:", unique(truth_population$scenario_no), "\n")
```

---

## Process Results Files

```{r process_results}
# Function to load and standardize a single RDS file
load_and_standardize <- function(file_path) {
  
  df <- readRDS(file_path)
  file_name <- basename(file_path)
  
  # Detect file format: new bonsaiforest2 models vs old format
  has_new_format <- "scenario_id" %in% names(df)
  
  if (has_new_format) {
    # NEW FORMAT: Global and OVAT models from bonsaiforest2
    # Create estimator from model_type and prior_name columns
    if ("model_type" %in% names(df) && "prior_name" %in% names(df)) {
      # Get the estimator name from data columns
      estimator_name <- df %>% 
        distinct(model_type, prior_name) %>% 
        pull() %>% 
        paste(collapse = "_")
      # Format: "{model_type}_{prior_name}"
      estimator_name <- paste(df$model_type[1], df$prior_name[1], sep = "_")
    } else {
      estimator_name <- str_remove(file_name, "\\.rds$")
    }
    
    cat("Processing:", estimator_name, "\n")
    
    # Find estimate, ci_lower, ci_upper columns (try multiple possible names)
    
    estimate_col <- names(df)[names(df) %in% c("estimate", "Effect", "Estimate", "trt_effect", "Median")] %>% 
      first(default = NA)
    lower_col <- names(df)[names(df) %in% c("lower", "ci_lower", "2.5%", "Lower", "CI_Lower")] %>% 
      first(default = NA)
    upper_col <- names(df)[names(df) %in% c("upper", "ci_upper", "97.5%", "Upper", "CI_Upper")] %>% 
      first(default = NA)
    
    if (is.na(estimate_col)) {
      cat("  WARNING: Could not find estimate column. Available columns:", 
          paste(names(df), collapse = ", "), "\n")
      return(tibble())
    }
    
    # Rename columns to standard names
    if (!is.na(lower_col) && lower_col != "ci_lower") {
      df <- df %>% rename(ci_lower = all_of(lower_col))
    } else if (is.na(lower_col)) {
      df$ci_lower <- NA_real_
    }
    
    if (!is.na(upper_col) && upper_col != "ci_upper") {
      df <- df %>% rename(ci_upper = all_of(upper_col))
    } else if (is.na(upper_col)) {
      df$ci_upper <- NA_real_
    }
    
    if (estimate_col != "estimate") {
      df <- df %>% rename(estimate = all_of(estimate_col))
    }
    
    df_clean <- df %>%
      mutate(
        scenario_no = as.character(scenario_id),
        estimator = estimator_name
      )
    
    # Select relevant columns
    cols_to_keep <- c("scenario_id", "replication_id", "scenario_no", "estimator", 
                      "estimate", "ci_lower", "ci_upper")
    # Add optional columns if they exist
    if ("level" %in% names(df_clean)) cols_to_keep <- c(cols_to_keep, "level")
    if ("subgroup_var" %in% names(df_clean)) cols_to_keep <- c(cols_to_keep, "subgroup_var")
    if ("covariate_tested" %in% names(df_clean)) cols_to_keep <- c(cols_to_keep, "covariate_tested")
    
    df_clean <- df_clean %>%
      dplyr::select(any_of(cols_to_keep))
    
  } else {
    # OLD FORMAT: naive/subgroup estimators from standard analysis
    estimator_name <- str_remove(file_name, "\\.rds$")
    
    cat("Processing:", estimator_name, "\n")
    
    # Try to find estimate column (handles various naming conventions)
    estimate_col <- names(df)[names(df) %in% c("estimate", "Effect", "Estimate", "trt_effect", "Median")] %>% 
      first(default = NA)
    se_col <- names(df)[names(df) %in% c("se", "SE", "trt_se", "std.error")] %>% 
      first(default = NA)
    replication_col <- names(df)[names(df) %in% c("sim_id", "simulation_id", "replication_id", "rep_id")] %>%
      first(default = NA)
    
    if (is.na(estimate_col)) {
      cat("  WARNING: Could not find estimate column in old format file. Available columns:", 
          paste(names(df), collapse = ", "), "\n")
      return(tibble())
    }
    
    if (is.na(replication_col)) {
      cat("  WARNING: Could not find replication ID column in old format file. Available columns:", 
          paste(names(df), collapse = ", "), "\n")
      return(tibble())
    }
    
    # Build base data frame
    df_clean <- df %>%
      mutate(
        scenario_id = as.integer(scenario_no),
        replication_id = as.integer(!!rlang::sym(replication_col)),
        estimator = estimator_name,
        estimate = !!rlang::sym(estimate_col),
        scenario_no = as.character(scenario_id)
      )
    
    # Add CIs if SE available
    if (!is.na(se_col)) {
      df_clean <- df_clean %>%
        mutate(
          ci_lower = estimate - 1.96 * !!rlang::sym(se_col),
          ci_upper = estimate + 1.96 * !!rlang::sym(se_col)
        )
    } else {
      df_clean <- df_clean %>%
        mutate(
          ci_lower = NA_real_,
          ci_upper = NA_real_
        )
    }
    
    # For subgroup estimator, keep subgroup_var and subgroup_level
    if ("subgroup_level" %in% names(df)) {
      df_clean <- df_clean %>%
        mutate(
          level = subgroup_level
        ) %>%
        dplyr::select(scenario_id, replication_id, scenario_no, estimator, estimate, 
               ci_lower, ci_upper, subgroup_var, level)
    } else {
      df_clean <- df_clean %>%
        dplyr::select(scenario_id, replication_id, scenario_no, estimator, estimate, 
               ci_lower, ci_upper)
    }
  }
  
  return(df_clean)
}

# Load all files
all_results <- map_dfr(all_files, load_and_standardize)

# Add scenario description from truth data
scenario_map <- simulation_truth %>%
  distinct(scenario_no, scenario) %>%
  mutate(scenario_no = as.character(scenario_no))

all_results <- all_results %>%
  left_join(scenario_map, by = "scenario_no")

cat("\nTotal rows loaded:", nrow(all_results), "\n")
cat("Estimators found:", paste(unique(all_results$estimator), collapse = ", "), "\n")
```

---

## Merge with Truth

```{r merge_truth}
# IMPORTANT NOTE ON TRUTH MATCHING:
# Results structure depends on model type:
#
# OLD FORMAT (naive/subgroup estimators):
# - Population estimator (Y~trt): Produces one estimate that applies to ALL subgroups.
#   We evaluate how well this single estimate captures the truth in each subgroup.
# - Subgroup estimator (Y~trt within group): Produces subgroup-specific estimates.
#   We evaluate each estimate against its corresponding subgroup truth.
#
# NEW FORMAT (global/OVAT shrinkage models from bonsaiforest2):
# - Global models: May produce population-level or subgroup-level estimates
# - OVAT models: Produce estimates for each covariate tested
#
# This reveals the bias-variance tradeoff:
# - In homogeneous scenarios: Global and population estimators have low RMSE
# - In heterogeneous scenarios: OVAT adapts and may have different RMSE profile

# Get subgroup-specific truth values
truth_subgroup <- simulation_truth %>%
  mutate(
    scenario_no = as.character(scenario_no),
    level = as.character(level)
  ) %>%
  dplyr::select(scenario_no, scenario, subgroup_var, level, trt_effect) %>%
  rename(truth_trt_effect = trt_effect)

# Merge results with truth
# For population estimators (no subgroup_var): Expand to all subgroups in each scenario
# For subgroup estimators (with subgroup_var): Match to corresponding subgroup truth
results_with_truth <- all_results %>%
  mutate(scenario_no = as.character(scenario_no)) %>%
  # Handle population estimators: expand to all subgroup truths
  {
    pop_results <- filter(., is.na(subgroup_var))
    subgroup_results <- filter(., !is.na(subgroup_var))
    
    # For population: join with all subgroup truths
    if (nrow(pop_results) > 0) {
      pop_joined <- pop_results %>%
        dplyr::select(-subgroup_var, -level) %>%
        left_join(truth_subgroup %>% dplyr::select(scenario_no, subgroup_var, level, truth_trt_effect),
                  by = "scenario_no", relationship = "many-to-many")
    } else {
      pop_joined <- pop_results %>% mutate(truth_trt_effect = NA_real_)
    }
    
    # For subgroup: join with matching subgroup truth
    if (nrow(subgroup_results) > 0) {
      subgroup_joined <- subgroup_results %>%
        left_join(truth_subgroup,
                  by = c("scenario_no", "subgroup_var", "level"))
    } else {
      subgroup_joined <- subgroup_results %>% mutate(truth_trt_effect = NA_real_)
    }
    
    bind_rows(pop_joined, subgroup_joined)
  }

cat("Results merged with truth values.\n")
cat("Total result-truth pairs:", nrow(results_with_truth), "\n")
cat("Pairs with truth values:", sum(!is.na(results_with_truth$truth_trt_effect)), "\n")

```

---

## Helper Functions

```{r helper_functions}
# Helper function to beautify estimator names
beautify_estimator_name <- function(estimator_name) {
  dplyr::case_when(
    # Old format estimators
    estimator_name == "naivepop" ~ "Population (Y~trt)",
    estimator_name == "subgroup" ~ "Subgroup (Y~trt within groups)",
    
    # New format: Global models with HN priors
    str_detect(estimator_name, "global.*phi_delta_plan_scenarios_1_6") ~ "Global - HN(δ_plan) [Sc 1-6]",
    str_detect(estimator_name, "global.*phi_delta_plan_scenarios_7_12") ~ "Global - HN(δ_plan) [Sc 7-12]",
    str_detect(estimator_name, "global.*phi_delta_half_scenarios_1_6") ~ "Global - HN(δ_plan/2) [Sc 1-6]",
    str_detect(estimator_name, "global.*phi_delta_half_scenarios_7_12") ~ "Global - HN(δ_plan/2) [Sc 7-12]",
    str_detect(estimator_name, "global.*phi_sigma_plan_scenarios_1_6") ~ "Global - HN(σ_plan) [Sc 1-6]",
    str_detect(estimator_name, "global.*phi_sigma_plan_scenarios_7_12") ~ "Global - HN(σ_plan) [Sc 7-12]",
    
    # New format: Horseshoe models
    str_detect(estimator_name, "horseshoe.*scenarios_1_6") ~ "Horseshoe (3 models) [Sc 1-6]",
    str_detect(estimator_name, "horseshoe.*scenarios_7_12") ~ "Horseshoe (3 models) [Sc 7-12]",
    
    # New format: OVAT models with HN priors
    str_detect(estimator_name, "ovat.*phi_sigma_plan_scenarios_1_6") ~ "OVAT - HN(σ_plan) [Sc 1-6]",
    str_detect(estimator_name, "ovat.*phi_sigma_plan_scenarios_7_12") ~ "OVAT - HN(σ_plan) [Sc 7-12]",
    str_detect(estimator_name, "ovat.*phi_delta_plan_scenarios_1_6") ~ "OVAT - HN(δ_plan) [Sc 1-6]",
    str_detect(estimator_name, "ovat.*phi_delta_plan_scenarios_7_12") ~ "OVAT - HN(δ_plan) [Sc 7-12]",
    str_detect(estimator_name, "ovat.*phi_delta_half_scenarios_1_6") ~ "OVAT - HN(δ_plan/2) [Sc 1-6]",
    str_detect(estimator_name, "ovat.*phi_delta_half_scenarios_7_12") ~ "OVAT - HN(δ_plan/2) [Sc 7-12]",
    
    # Legacy cases
    estimator_name == "Global_full_hierarchical_05" ~ "GM - HN(0.5)",
    estimator_name == "Global_full_hierarchical" ~ "GM - HN(1)",
    estimator_name == "Global_horseshoe_1" ~ "GM - Horseshoe(1)",
    estimator_name == "OVAT_halfnormal_05" ~ "OVAT - HN(0.5)",
    estimator_name == "OVAT_halfnormal" ~ "OVAT - HN(1)",
    
    TRUE ~ estimator_name
  )
}

# Helper function to format numbers
display <- function(x, digits = 3) {
  formatC(x, format = "f", digits = digits)
}

# Helper function to format percentages
getPercentage <- function(x, digits = 0) {
  val <- 100 * x
  ifelse(is.na(val), "NA",
    paste(formatC(val, digits = digits, format = "f"), "%", sep = "")
  )
}
```

---

## Calculate Performance Metrics

```{r calculate_metrics}
# Prepare data for metrics calculation
# Both estimators compared to SUBGROUP-SPECIFIC truths:
# - Population estimator: Single estimate compared to all subgroup truths
#   (High RMSE in heterogeneous scenarios as diversity in truths increases)
# - Subgroup estimator: Each subgroup estimate compared to its specific truth
#   (Adapts to heterogeneity, lower RMSE in heterogeneous scenarios)

results_for_metrics <- results_with_truth %>%
  filter(!is.na(truth_trt_effect)) %>%
  mutate(
    error = estimate - truth_trt_effect,
    ci_contains_truth = (truth_trt_effect >= ci_lower) & (truth_trt_effect <= ci_upper)
  )

# Calculate RMSE, Bias, and Coverage by estimator and scenario
performance_by_scenario <- results_for_metrics %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    n_estimates = n(),
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    Bias = mean(error, na.rm = TRUE),
    AbsBias = mean(abs(error), na.rm = TRUE),
    Coverage = mean(ci_contains_truth, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  left_join(
    simulation_truth %>% 
      distinct(scenario_no, scenario) %>%
      mutate(scenario_no = as.character(scenario_no)),
    by = "scenario_no"
  ) %>%
  arrange(as.numeric(scenario_no), estimator)

# Calculate overall metrics (across all scenarios)
performance_overall <- results_for_metrics %>%
  group_by(estimator) %>%
  summarise(
    n_estimates = n(),
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    Bias = mean(error, na.rm = TRUE),
    AbsBias = abs(Bias),
    Coverage = mean(ci_contains_truth, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(estimator)

cat("Performance metrics calculated.\n")
cat("Both estimators evaluated against subgroup-specific truth values.\n")
cat("Population estimator: Single estimate repeated across all subgroups for comparison.\n")
```

---

## Bias-Variance Diagnostics

```{r bias_variance_check}
# Key diagnostic: Check if subgroup estimator is UNBIASED and population is BIASED
# 
# Theory:
# - Subgroup estimator (Y~trt within group): Should be UNBIASED (Bias ≈ 0) with HIGH VARIANCE
# - Population estimator (Y~trt overall): Could be BIASED in heterogeneous scenarios with LOW VARIANCE
#
# RMSE = sqrt(Bias² + Variance), so:
# - Subgroup RMSE ≈ sqrt(Variance) since Bias ≈ 0
# - Population RMSE ≈ sqrt(Bias² + Variance)

diagnostics <- results_for_metrics %>%
  group_by(scenario_no, scenario, estimator) %>%
  summarise(
    n = n(),
    mean_error = mean(error, na.rm = TRUE),
    mean_abs_error = mean(abs(error), na.rm = TRUE),
    variance_error = var(error, na.rm = TRUE),
    bias_squared = mean_error^2,
    rmse = sqrt(mean(error^2, na.rm = TRUE)),
    .groups = 'drop'
  ) %>%
  mutate(
    bias_component = sqrt(bias_squared),
    variance_component = sqrt(variance_error)
  ) %>%
  arrange(as.numeric(scenario_no), estimator)


```



```{r}
diagnostics_2 <- results_for_metrics %>%
  filter(estimator %in% c("continuous_naivepop", "continuous_subgroup")) %>%
  group_by(scenario_no, scenario, estimator, subgroup_var, level) %>%
  summarise(
    n = n(),
    mean_estimate = mean(estimate, na.rm = TRUE),
    mean_bias = mean(error, na.rm = TRUE),
    mean_abs_bias = mean(abs(error), na.rm = TRUE),
    rmse = sqrt(mean(error^2, na.rm = TRUE)),
    .groups = 'drop'
  )
```


---

## Results: Overall Performance

```{r table_overall}
performance_overall %>%
  mutate(
    Estimator = beautify_estimator_name(estimator),
    RMSE = round(RMSE, 4),
    Bias = round(Bias, 4),
    AbsBias = round(AbsBias, 4),
    Coverage = paste0(round(Coverage * 100, 1), "%")
  ) %>%
  dplyr::select(Estimator, RMSE, Bias, AbsBias, Coverage) %>%
  kable(caption = "Overall Performance Metrics (All Scenarios Combined)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

---

## Results: Performance by Scenario

```{r table_by_scenario}
performance_by_scenario %>%
  filter(estimator %in% c('continuous_naivepop', "continuous_subgroup"))%>%
  mutate(
    Estimator = beautify_estimator_name(estimator),
    RMSE = round(RMSE, 4),
    Bias = round(Bias, 4),
    AbsBias = round(AbsBias, 4),
    Coverage = paste0(round(Coverage * 100, 1), "%")
  ) %>%
  dplyr::select(Scenario = scenario_no, Scenario_Name = scenario, 
         Estimator, RMSE, Bias, AbsBias, Coverage) %>%
  kable(caption = "Performance Metrics by Scenario") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

---

## Visualization: RMSE by Scenario Group

```{r plot_rmse_grouped, fig.height=10}
# Create scenario grouping
scenario_groups <- data.frame(
  scenario_no = as.character(1:12),
  group = rep(c("Group 1\n(Scenarios 1-3)", 
                "Group 2\n(Scenarios 4-6)", 
                "Group 3\n(Scenarios 7-9)",
                "Group 4\n(Scenarios 10-12)"), each = 3)
)

# Create plot data with scenario groups and clean estimator names
plot_data_rmse <- performance_by_scenario %>%
  mutate(
    scenario_no_numeric = as.numeric(scenario_no)
  ) %>%
  left_join(scenario_groups, by = "scenario_no") %>%
  mutate(
    estimator = beautify_estimator_name(estimator),
    group = factor(group, levels = c("Group 1\n(Scenarios 1-3)", 
                                     "Group 2\n(Scenarios 4-6)", 
                                     "Group 3\n(Scenarios 7-9)",
                                     "Group 4\n(Scenarios 10-12)"))
  )

# RMSE plot with grouping
ggplot(plot_data_rmse, aes(x = scenario_no_numeric, y = RMSE, 
                            color = estimator, group = estimator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  facet_wrap(~group, scales = "free_x", nrow = 2) +
  labs(
    title = "RMSE by Scenario Group and Estimator",
    subtitle = "Root Mean Squared Error - Treatment Effect Scale",
    x = "Scenario",
    y = "RMSE",
    color = "Estimator"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold", size = 11)
  ) +
  guides(color = guide_legend(ncol = 2))
```

---

## Visualization: Overall RMSE Trend

```{r plot_rmse_overall, fig.height=7}
# Overall RMSE trend across all scenarios
ggplot(plot_data_rmse, aes(x = scenario_no_numeric, y = RMSE, 
                            color = estimator, group = estimator)) +
  geom_line(size = 1.3) +
  geom_point(size = 3.5) +
  scale_x_continuous(breaks = 1:12) +
  labs(
    title = "RMSE Across All Scenarios",
    subtitle = "Treatment Effect Scale",
    x = "Scenario",
    y = "RMSE",
    color = "Estimator"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 2))
```

---





---

## Heatmap: RMSE Comparison

```{r heatmap_rmse, fig.height=8}
# Create heatmap
ggplot(plot_data_rmse, aes(x = scenario_no_numeric, y = estimator, fill = RMSE)) +
  geom_tile(color = "white", size = 0.5) +
  geom_text(aes(label = round(RMSE, 3)), size = 3) +
  scale_fill_gradient2(
    low = "#2166ac", 
    mid = "#f7f7f7", 
    high = "#b2182b",
    midpoint = median(plot_data_rmse$RMSE, na.rm = TRUE),
    name = "RMSE"
  ) +
  scale_x_continuous(breaks = 1:12, labels = 1:12) +
  labs(
    title = "RMSE Heatmap: Estimator × Scenario",
    subtitle = "Lower values (blue) indicate better performance",
    x = "Scenario",
    y = "Estimator"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    axis.text.y = element_text(size = 9),
    legend.position = "right"
  )
```

---

## Heatmap: Absolute Bias Comparison

```{r heatmap_abs_bias, fig.height=8}
# Create absolute bias heatmap
ggplot(plot_data_rmse, aes(x = scenario_no_numeric, y = estimator, fill = AbsBias)) +
  geom_tile(color = "white", size = 0.5) +
  geom_text(aes(label = round(AbsBias, 3)), size = 3) +
  scale_fill_gradient2(
    low = "#2166ac", 
    mid = "#f7f7f7", 
    high = "#b2182b",
    midpoint = median(plot_data_rmse$AbsBias, na.rm = TRUE),
    name = "|Bias|"
  ) +
  scale_x_continuous(breaks = 1:12, labels = 1:12) +
  labs(
    title = "Absolute Bias Heatmap: Estimator × Scenario",
    subtitle = "Lower values (blue) indicate less deviation from truth",
    x = "Scenario",
    y = "Estimator"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    axis.text.y = element_text(size = 9),
    legend.position = "right"
  )
```

---

## Heatmap: Coverage Comparison

```{r heatmap_coverage, fig.height=8}
# Create coverage heatmap
ggplot(plot_data_rmse, aes(x = scenario_no_numeric, y = estimator, fill = Coverage)) +
  geom_tile(color = "white", size = 0.5) +
  geom_text(aes(label = paste0(round(Coverage * 100, 1), "%")), size = 3) +
  scale_fill_gradient2(
    low = "#b2182b", 
    mid = "#f7f7f7", 
    high = "#2166ac",
    midpoint = 0.95,
    limits = c(0.8, 1.0),
    name = "Coverage"
  ) +
  scale_x_continuous(breaks = 1:12, labels = 1:12) +
  labs(
    title = "Coverage Heatmap: Estimator × Scenario",
    subtitle = "Target coverage = 95% (blue indicates good coverage)",
    x = "Scenario",
    y = "Estimator"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    axis.text.y = element_text(size = 9),
    legend.position = "right"
  )
```

---

## Summary Statistics by Scenario Group

```{r performance_by_group}
# Add group assignments and calculate summary by group
plot_data_rmse_summary <- plot_data_rmse %>%
  group_by(group, estimator) %>%
  summarise(
    mean_RMSE = mean(RMSE, na.rm = TRUE),
    sd_RMSE = sd(RMSE, na.rm = TRUE),
    min_RMSE = min(RMSE, na.rm = TRUE),
    max_RMSE = max(RMSE, na.rm = TRUE),
    mean_Bias = mean(Bias, na.rm = TRUE),
    mean_AbsBias = mean(AbsBias, na.rm = TRUE),
    mean_Coverage = mean(Coverage, na.rm = TRUE),
    .groups = 'drop'
  )

plot_data_rmse_summary %>%
  mutate(
    `Mean RMSE` = display(mean_RMSE, 4),
    `Range RMSE` = paste0(display(min_RMSE, 4), " - ", display(max_RMSE, 4)),
    `Mean Bias` = display(mean_Bias, 4),
    `Mean |Bias|` = display(mean_AbsBias, 4),
    `Mean Coverage` = getPercentage(mean_Coverage, 1)
  ) %>%
  dplyr::select(Group = group, Estimator = estimator, 
         `Mean RMSE`, `Range RMSE`, `Mean Bias`, `Mean |Bias|`, `Mean Coverage`) %>%
  kable(caption = "Performance Summary by Scenario Group") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

## Notes on Scenarios

```{r scenario_descriptions}
# Define scenario descriptions for continuous outcomes
scenario_info <- tibble(
  Scenario = 1:12,
  Description = c(
    "1: Homogeneous effect (no treatment heterogeneity)",
    "2: Mild heterogeneity by X11cat",
    "3: Strong heterogeneity by X11cat",
    "4: Homogeneous effect variant",
    "5: Mild heterogeneity variant",
    "6: Strong heterogeneity variant",
    "7: Homogeneous with different effect size",
    "8: Mild heterogeneity with different effect size",
    "9: Strong heterogeneity with different effect size",
    "10: Homogeneous with another variant",
    "11: Mild heterogeneity with another variant",
    "12: Strong heterogeneity with another variant"
  )
)

scenario_info %>%
  kable(caption = "Scenario Descriptions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

---

**Document generated**: `r Sys.time()`
