---
title: "RHS Global Methods - Simulation Results Analysis"
subtitle: "RMSE, Bias, Coverage, and Extreme Subgroup Identification Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.width = 12, fig.height = 8)

library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(ggplot2)
library(knitr)
library(kableExtra)
```

# RHS Global Methods - Simulation Results Analysis

This document analyzes Regularized Horseshoe (RHS) global method simulation results, calculating:
- **RMSE** (Root Mean Squared Error)
- **Bias** (Mean difference between estimate and truth)
- **Coverage** (Proportion of 95% CIs containing the true value)
- **Extreme Subgroup Identification Accuracy** (Proportion of replications correctly identifying the subgroup most different from the overall treatment effect)

All metrics are calculated on the log-scale (log hazard ratios).

---

## Load Results

```{r load_results}
# Set paths
base_path <- "/home/pedreram/bonsaiforest2/simulations/TTE"
results_path <- file.path(base_path, "Results")

# Find all RHS RDS files
rhs_pattern <- "^tte_RHS.*\\.rds$"
rhs_files <- list.files(results_path, pattern = rhs_pattern, full.names = TRUE)

# Sort files by name for consistent ordering
rhs_files <- sort(rhs_files)

cat("Found", length(rhs_files), "RHS result files:\n")
cat(paste("-", basename(rhs_files), collapse = "\n"), "\n")

# Check dimensions of each RDS file
cat("\n--- Dimensions of Each RHS RDS File ---\n")
cat(strrep("-", 80), "\n")
cat(sprintf("%-50s %10s %10s", "File Name", "Rows", "Columns"), "\n")
cat(strrep("-", 80), "\n")

file_summary <- data.frame()

for (file_path in rhs_files) {
  df <- readRDS(file_path)
  file_name <- basename(file_path)
  n_rows <- nrow(df)
  n_cols <- ncol(df)
  
  cat(sprintf("%-50s %10s %10s\n", file_name, n_rows, n_cols))
  
  # Store summary
  file_summary <- rbind(file_summary, data.frame(
    file = file_name,
    rows = n_rows,
    cols = n_cols,
    stringsAsFactors = FALSE
  ))
  
  rm(df)
}

cat(strrep("-", 80), "\n")
cat("TOTAL ROWS:", sum(file_summary$rows), "\n")
cat("TOTAL FILES:", nrow(file_summary), "\n")
cat("\n")

# Show summary table
kable(file_summary, caption = "Dimensions of Each RHS RDS File") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

---

## Load Truth Values

```{r load_truth}
# Load truth data
truth_file <- file.path(base_path, "Scenarios", "truth.RData")
load(truth_file)
cat("Loaded truth values from:", truth_file, "\n")

# Extract subgroup-level truth (log hazard ratios)
truth_subgroup <- simul_parameter$true_subgroup_ahr %>%
  mutate(scenario_no = as.character(row_number())) %>%
  pivot_longer(
    cols = -scenario_no,
    names_to = "subgroup_raw",
    values_to = "truth_ahr"
  ) %>%
  mutate(
    # Clean subgroup names to match results (remove x_, dots, etc.)
    join_key = gsub("x_|\\.", "", subgroup_raw) %>% tolower(),
    truth_log = log(truth_ahr)
  ) %>%
  dplyr::select(scenario_no, join_key, truth_log, truth_ahr)

# Extract population-level truth
truth_population <- simul_parameter$true_overall_results %>%
  mutate(
    scenario_no = as.character(1:6),
    truth_pop_ahr = AHR,
    truth_pop_log = log(AHR)
  ) %>%
  dplyr::select(scenario_no, truth_pop_ahr, truth_pop_log)

cat("\nTruth data loaded successfully.\n")
cat("Scenarios:", unique(truth_subgroup$scenario_no), "\n")
cat("Subgroups per scenario:", length(unique(truth_subgroup$join_key)), "\n")

# Check dimensions of loaded truth dataframes
cat("\n--- Truth Dataframe Dimensions ---\n")
cat("truth_subgroup: ", nrow(truth_subgroup), " rows × ", ncol(truth_subgroup), " columns\n", sep = "")
cat("truth_population: ", nrow(truth_population), " rows × ", ncol(truth_population), " columns\n", sep = "")
```

---

## Process RHS Results Files

```{r process_results}
# Function to load and standardize a single RHS RDS file
load_and_standardize_rhs <- function(file_path) {
  
  df <- readRDS(file_path)
  
  # RHS estimators from bonsaiforest2
  df_clean <- df %>%
    filter(Subgroup != "Overall") %>%
    mutate(
      scenario_id = as.integer(scenario_id),
      replication_id = as.integer(replication_id),
      # Clean subgroup names (e.g., "x_1: a" -> "1a")
      join_key = gsub("x_|: | ", "", Subgroup) %>% 
                 str_replace_all("[\\._\\s]", "") %>% 
                 tolower(),
      estimator = paste(model_type, prior_name, sep = "_"),
      # Convert from hazard ratio scale to log scale
      estimate_log = log(Median),
      ci_lower = log(CI_Lower),
      ci_upper = log(CI_Upper)
    )
  
  # Select common columns
  df_clean %>%
    dplyr::select(scenario_id, replication_id, estimator, join_key, 
           estimate_log, ci_lower, ci_upper) %>%
    mutate(scenario_no = as.character(scenario_id))
}

# Load all RHS files and track dimensions
cat("--- Processing Each RHS RDS File ---\n")
cat(strrep("-", 100), "\n")
cat(sprintf("%-50s %10s %10s %25s", "File Name", "Rows (raw)", "Rows (clean)", "Estimator"), "\n")
cat(strrep("-", 100), "\n")

all_results_list <- list()

for (i in seq_along(rhs_files)) {
  file_path <- rhs_files[i]
  file_name <- basename(file_path)
  
  # Load raw
  df_raw <- readRDS(file_path)
  n_rows_raw <- nrow(df_raw)
  
  # Process
  df_clean <- load_and_standardize_rhs(file_path)
  n_rows_clean <- nrow(df_clean)
  estimator <- unique(df_clean$estimator)[1]
  
  cat(sprintf("%-50s %10d %10d %25s\n", file_name, n_rows_raw, n_rows_clean, estimator))
  
  all_results_list[[i]] <- df_clean
}

cat(strrep("-", 100), "\n")

# Combine all results
all_results <- bind_rows(all_results_list)

cat("\nTotal rows after combining:", nrow(all_results), "\n")
cat("RHS Estimators found:", paste(unique(all_results$estimator), collapse = ", "), "\n")

# Check dimensions before merging
cat("\n--- Dataframe Dimensions Before Merging ---\n")
cat("all_results: ", nrow(all_results), " rows × ", ncol(all_results), " columns\n", sep = "")
cat("  Unique scenarios: ", n_distinct(all_results$scenario_no), "\n", sep = "")
cat("  Unique estimators: ", n_distinct(all_results$estimator), "\n", sep = "")
cat("  Unique subgroups: ", n_distinct(all_results$join_key), "\n", sep = "")
```

---

## Merge with Truth

```{r merge_truth}
# Show dimensions before merge
cat("\n--- Before Merging ---\n")
cat("all_results: ", nrow(all_results), " rows\n", sep = "")
cat("truth_subgroup: ", nrow(truth_subgroup), " rows\n", sep = "")
cat("truth_population: ", nrow(truth_population), " rows\n", sep = "")

# Merge subgroup-level truth
results_with_truth <- all_results %>%
  left_join(truth_subgroup, by = c("scenario_no", "join_key")) %>%
  left_join(truth_population, by = "scenario_no")

cat("\n--- After Merging ---\n")
cat("results_with_truth: ", nrow(results_with_truth), " rows × ", ncol(results_with_truth), " columns\n", sep = "")

# Check for missing truth values
missing_truth <- results_with_truth %>%
  filter(is.na(truth_log)) %>%
  distinct(scenario_no, join_key)

if (nrow(missing_truth) > 0) {
  cat("\nWarning: Missing truth values for:\n")
  print(missing_truth)
} else {
  cat("\nAll estimates successfully matched with truth values.\n")
}

cat("Total estimates with truth:", sum(!is.na(results_with_truth$truth_log)), "\n")
```

---

## Helper Functions

```{r helper_functions}
# Helper function to beautify RHS estimator names
beautify_rhs_estimator <- function(estimator_name) {
  dplyr::case_when(
    estimator_name == "Global_RHS_theta0_1_s_2" ~ "RHS (θ₀=1, σ=2)",
    estimator_name == "Global_RHS_theta0_delta_plan_half_s_2" ~ "RHS (θ₀=δ/2, σ=2)",
    estimator_name == "Global_RHS_theta0_delta_plan_half_s_delta_plan" ~ "RHS (θ₀=δ/2, σ=δ)",
    estimator_name == "Global_RHS_theta0_delta_plan_s_2" ~ "RHS (θ₀=δ, σ=2)",
    estimator_name == "Global_RHS_theta0_delta_plan_s_2delta_plan" ~ "RHS (θ₀=δ, σ=2δ)",
    TRUE ~ estimator_name
  )
}

# Helper function to format numbers
display <- function(x, digits = 2) {
  formatC(x, format = "f", digits = digits)
}

# Helper function to format percentages
getPercentage <- function(x, digits = 0) {
  val <- 100 * x
  ifelse(is.na(val), "NA",
    paste(formatC(val, digits = digits, format = "f"), "%", sep = "")
  )
}
```

---

## Calculate Performance Metrics

```{r calculate_metrics}
# Calculate RMSE, Bias, and Coverage by estimator and scenario
performance_by_scenario <- results_with_truth %>%
  filter(!is.na(truth_log)) %>%
  mutate(
    # Calculate error and coverage indicator
    error = estimate_log - truth_log,
    ci_contains_truth = (truth_log >= ci_lower) & (truth_log <= ci_upper)
  ) %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    n_estimates = n(),
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    Bias = mean(error, na.rm = TRUE),
    Coverage = mean(ci_contains_truth, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(scenario_no, estimator)

# Calculate overall metrics (across all scenarios)
performance_overall <- results_with_truth %>%
  filter(!is.na(truth_log)) %>%
  mutate(
    error = estimate_log - truth_log,
    ci_contains_truth = (truth_log >= ci_lower) & (truth_log <= ci_upper)
  ) %>%
  group_by(estimator) %>%
  summarise(
    n_estimates = n(),
    RMSE = sqrt(mean(error^2, na.rm = TRUE)),
    Bias = mean(error, na.rm = TRUE),
    Coverage = mean(ci_contains_truth, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(estimator)

cat("RHS performance metrics calculated.\n")
```

---

## Visualization: RMSE by Scenario

```{r plot_rmse, fig.height=7}
# Define scenario labels
scenario_labels <- c(
  `1` = "Scenario 1:\nPositive (homogeneous)",
  `2` = "Scenario 2:\nPositive (except 1)",
  `3` = "Scenario 3:\nNegative (except 1)",
  `4` = "Scenario 4:\nMild heterogeneity",
  `5` = "Scenario 5:\nStrong heterogeneity",
  `6` = "Scenario 6:\nMisspecified"
)

# Create plot data with scenario labels and clean estimator names
plot_data_rmse <- performance_by_scenario %>%
  mutate(
    scenario_no = as.numeric(scenario_no),
    scenario_label = scenario_labels[as.character(scenario_no)],
    estimator = beautify_rhs_estimator(estimator)
  ) 

# RMSE plot
ggplot(plot_data_rmse, aes(x = scenario_no, y = RMSE, 
                            color = estimator, group = estimator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "RMSE by Scenario - RHS Global Methods",
    subtitle = "Root Mean Squared Error on log-hazard ratio scale",
    x = "Scenario",
    y = "RMSE",
    color = "RHS Model"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 2))
```

---

## Visualization: RMSE - Fixed σ=2 Estimators

```{r plot_rmse_sigma_fixed, fig.height=7}
# Plot with estimators that have σ=2 fixed
plot_data_sigma_fixed <- performance_by_scenario %>%
  filter(estimator %in% c(
    "Global_RHS_theta0_1_s_2",
    "Global_RHS_theta0_delta_plan_half_s_2",
    "Global_RHS_theta0_delta_plan_s_2"
  )) %>%
  mutate(
    scenario_no = as.numeric(scenario_no),
    scenario_label = scenario_labels[as.character(scenario_no)],
    estimator = beautify_rhs_estimator(estimator)
  )

# RMSE plot - σ=2 fixed
ggplot(plot_data_sigma_fixed, aes(x = scenario_no, y = RMSE, 
                                   color = estimator, group = estimator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "RMSE by Scenario - RHS Global Methods with σ=2 (Fixed)",
    subtitle = "Root Mean Squared Error on log-hazard ratio scale",
    x = "Scenario",
    y = "RMSE",
    color = "RHS Model"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 1))
```

---

## Visualization: RMSE - Changing σ with Delta Estimators

```{r plot_rmse_sigma_changing, fig.height=7}
# Plot with estimators that have σ changing with delta
# Note: θ₀=1, σ=2 case is included here as delta_plan=1 makes σ=2 equivalent to σ=δ
plot_data_sigma_changing <- performance_by_scenario %>%
  filter(estimator %in% c(
    "Global_RHS_theta0_1_s_2",
    "Global_RHS_theta0_delta_plan_half_s_delta_plan",
    "Global_RHS_theta0_delta_plan_s_2delta_plan"
  )) %>%
  mutate(
    scenario_no = as.numeric(scenario_no),
    scenario_label = scenario_labels[as.character(scenario_no)],
    estimator = beautify_rhs_estimator(estimator)
  )

# RMSE plot - σ changing
ggplot(plot_data_sigma_changing, aes(x = scenario_no, y = RMSE, 
                                      color = estimator, group = estimator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "RMSE by Scenario - RHS Global Methods with σ Changing (σ ∝ δ)",
    subtitle = "Root Mean Squared Error on log-hazard ratio scale; θ₀=1,σ=2 included (δ=1 case)",
    x = "Scenario",
    y = "RMSE",
    color = "RHS Model"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 1))
```

## weird stuff
## Visualization: RMSE - Fixed σ=2 Estimators

```{r plot_rmse_sigma_fixed, fig.height=7}
# Plot with estimators that have σ=2 fixed
plot_data_sigma_fixed <- performance_by_scenario %>%
  filter(estimator %in% c(
    "Global_RHS_theta0_delta_plan_s_2",
    "Global_RHS_theta0_delta_plan_s_2delta_plan"
  )) %>%
  mutate(
    scenario_no = as.numeric(scenario_no),
    scenario_label = scenario_labels[as.character(scenario_no)],
    estimator = beautify_rhs_estimator(estimator)
  )

# RMSE plot - σ=2 fixed
ggplot(plot_data_sigma_fixed, aes(x = scenario_no, y = RMSE, 
                                   color = estimator, group = estimator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "RMSE by Scenario - RHS Global Methods with θ₀=δ (Fixed)",
    subtitle = "Root Mean Squared Error on log-hazard ratio scale",
    x = "Scenario",
    y = "RMSE",
    color = "RHS Model"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 1))
```

## weird stuff
## Visualization: RMSE - Fixed σ=2 Estimators

```{r plot_rmse_sigma_fixed, fig.height=7}
# Plot with estimators that have σ=2 fixed
plot_data_sigma_fixed <- performance_by_scenario %>%
  filter(estimator %in% c(
    "Global_RHS_theta0_delta_plan_half_s_2",
    "Global_RHS_theta0_delta_plan_half_s_delta_plan"
  )) %>%
  mutate(
    scenario_no = as.numeric(scenario_no),
    scenario_label = scenario_labels[as.character(scenario_no)],
    estimator = beautify_rhs_estimator(estimator)
  )

# RMSE plot - σ=2 fixed
ggplot(plot_data_sigma_fixed, aes(x = scenario_no, y = RMSE, 
                                   color = estimator, group = estimator)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "RMSE by Scenario - RHS Global Methods with θ₀=δ/2 (Fixed)",
    subtitle = "Root Mean Squared Error on log-hazard ratio scale",
    x = "Scenario",
    y = "RMSE",
    color = "RHS Model"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(ncol = 1))
```
---

## Heatmap: RMSE Comparison

```{r heatmap_rmse, fig.height=8}
# Create heatmap
ggplot(plot_data_rmse, aes(x = scenario_no, y = estimator, fill = RMSE)) +
  geom_tile(color = "white", size = 0.5) +
  geom_text(aes(label = round(RMSE, 3)), size = 3.5) +
  scale_fill_gradient2(
    low = "white",
    mid = "yellow",
    high = "red",
    midpoint = median(plot_data_rmse$RMSE, na.rm = TRUE),
    name = "RMSE"
  ) +
  scale_x_continuous(breaks = 1:6, labels = 1:6) +
  labs(
    title = "RMSE Heatmap: RHS Model × Scenario",
    subtitle = "Lower values (white) indicate better performance",
    x = "Scenario",
    y = "RHS Model"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.text.x = element_text(angle = 0, hjust = 0.5, size = 9),
    axis.text.y = element_text(size = 9),
    legend.position = "right"
  )
```

---

## Performance Table

```{r performance_table}
# Create subgroup-level performance metrics
subgroup_performance <- results_with_truth %>%
  filter(!is.na(truth_log)) %>%
  group_by(scenario_no, join_key, estimator) %>%
  summarize(
    truth_log = first(truth_log),
    truth_pop_log = first(truth_pop_log),
    bias = mean(estimate_log - truth_log, na.rm = TRUE),
    RMSE = sqrt(mean((estimate_log - truth_log)^2, na.rm = TRUE)),
    coverage = mean((truth_log >= ci_lower) & (truth_log <= ci_upper), na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  mutate(
    estimator_display = beautify_rhs_estimator(estimator),
    # For TTE, use log(1.1) as threshold for heterogeneity
    threshold = log(1.1),
    heterogeneous = (abs(truth_log - truth_pop_log) > threshold)
  ) %>%
  dplyr::select(-threshold)

# Summary by estimator (overall)
sum_RMSEbias <- subgroup_performance %>%
  group_by(estimator_display) %>%
  summarize(
    RMSE = paste(display(mean(RMSE, na.rm = TRUE), 2), " (", 
                 display(min(RMSE, na.rm = TRUE), 2), "-", 
                 display(max(RMSE, na.rm = TRUE), 2), ")", sep = ""),
    abs_bias = paste(display(mean(abs(bias), na.rm = TRUE), 2), " (", 
                     display(min(abs(bias), na.rm = TRUE), 2), "-", 
                     display(max(abs(bias), na.rm = TRUE), 2), ")", sep = ""),
    .groups = 'drop'
  ) %>%
  arrange(estimator_display)

sum_coverage <- subgroup_performance %>%
  group_by(estimator_display) %>%
  summarize(
    coverage = ifelse(
      all(is.na(coverage)), "NA",
      paste(getPercentage(mean(coverage, na.rm = TRUE)), " (", 
            getPercentage(min(coverage, na.rm = TRUE)), "-", 
            getPercentage(max(coverage, na.rm = TRUE)), ")", sep = "")
    ),
    .groups = 'drop'
  ) %>%
  arrange(estimator_display)

# Summary by heterogeneity
sum_RMSEbias_het <- subgroup_performance %>%
  group_by(heterogeneous, estimator_display) %>%
  summarize(
    RMSE = paste(display(mean(RMSE, na.rm = TRUE), 2), " (", 
                 display(min(RMSE, na.rm = TRUE), 2), "-", 
                 display(max(RMSE, na.rm = TRUE), 2), ")", sep = ""),
    abs_bias = paste(display(mean(abs(bias), na.rm = TRUE), 2), " (", 
                     display(min(abs(bias), na.rm = TRUE), 2), "-", 
                     display(max(abs(bias), na.rm = TRUE), 2), ")", sep = ""),
    .groups = 'drop'
  ) %>%
  arrange(estimator_display)

sum_coverage_het <- subgroup_performance %>%
  group_by(heterogeneous, estimator_display) %>%
  summarize(
    coverage = ifelse(
      all(is.na(coverage)), "NA",
      paste(getPercentage(mean(coverage, na.rm = TRUE)), " (", 
            getPercentage(min(coverage, na.rm = TRUE)), "-", 
            getPercentage(max(coverage, na.rm = TRUE)), ")", sep = "")
    ),
    .groups = 'drop'
  ) %>%
  arrange(estimator_display)

# Assemble table
estimator_names <- sort(unique(subgroup_performance$estimator_display))
estimator_rows <- paste("  ", estimator_names, sep = "")
n_estimators <- length(estimator_names)

estimators_with_coverage <- sum_coverage$estimator_display[sum_coverage$coverage != "NA"]
estimator_rows_coverage <- paste("  ", estimators_with_coverage, sep = "")
n_estimators_coverage <- length(estimators_with_coverage)

criteria_vec <- c(
  "RMSE", estimator_rows,
  "Absolute bias", estimator_rows,
  "Coverage of 95% CI", estimator_rows_coverage
)

overall_vec <- c(
  " ", sum_RMSEbias$RMSE,
  " ", sum_RMSEbias$abs_bias,
  " ", sum_coverage$coverage[sum_coverage$coverage != "NA"]
)

homo_vec <- c(
  " ", with(sum_RMSEbias_het, RMSE[!heterogeneous]),
  " ", with(sum_RMSEbias_het, abs_bias[!heterogeneous]),
  " ", with(sum_coverage_het, coverage[!heterogeneous & coverage != "NA"])
)

het_vec <- c(
  " ", with(sum_RMSEbias_het, RMSE[heterogeneous]),
  " ", with(sum_RMSEbias_het, abs_bias[heterogeneous]),
  " ", with(sum_coverage_het, coverage[heterogeneous & coverage != "NA"])
)

pad_to <- length(criteria_vec)
pad <- function(x, n) {
  x <- as.character(x)
  if (length(x) < n) length(x) <- n
  x[is.na(x)] <- ""
  x
}

performance_table <- data.frame(
  Criterion = criteria_vec,
  All_Subgroups = pad(overall_vec, pad_to),
  Homogeneous = pad(homo_vec, pad_to),
  Heterogeneous = pad(het_vec, pad_to),
  stringsAsFactors = FALSE
)

# Calculate footnotes
n_scenarios <- n_distinct(subgroup_performance$scenario_no)
n_subgroups_per_scenario <- n_distinct(subgroup_performance$join_key)
total_subgroups <- n_scenarios * n_subgroups_per_scenario

count_het <- subgroup_performance %>%
  filter(estimator == "Global_RHS_theta0_1_s_2") %>%
  summarise(total_het = sum(heterogeneous))
total_het <- sum(count_het$total_het)
total_homo <- total_subgroups - total_het

# Display table
kbl(performance_table,
    caption = "Performance Metrics for RHS Global Methods",
    booktabs = TRUE,
    linesep = c(
      "", rep("", n_estimators),
      "\\addlinespace", rep("", n_estimators),
      "\\addlinespace", rep("", n_estimators_coverage)
    ),
    col.names = c("Criterion", "All subgroups", "Homogeneous", "Heterogeneous")
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE) %>%
  footnote(
    general = c(
      paste("Summary across", total_subgroups, "subgroups (", n_subgroups_per_scenario, 
            "subgroups ×", n_scenarios, "scenarios) including", total_homo, 
            "homogeneous and", total_het, "heterogeneous subgroups.")
    ),
    general_title = ""
  )
```

---

## Identify True Worst Subgroup

```{r identify_worst_subgroup}
# For each scenario, identify the subgroup with the worst (lowest) treatment effect
true_worst_subgroup <- truth_subgroup %>%
  group_by(scenario_no) %>%
  mutate(
    is_worst = truth_ahr == max(truth_ahr, na.rm = TRUE)
  ) %>%
  filter(is_worst) %>%
  dplyr::select(scenario_no, join_key, truth_ahr, truth_log) %>%
  rename(true_worst_subgroup = join_key, true_worst_ahr = truth_ahr, true_worst_effect = truth_log) %>%
  distinct()

cat("\nTrue Worst Subgroup by Scenario (Lowest Treatment Effect):\n")
print(true_worst_subgroup)
```

---

## Worst Treatment Effect Identification Accuracy

```{r worst_accuracy}
# For each scenario, replication, and estimator, identify the subgroup with the worst (lowest) estimate
predicted_worst <- results_with_truth %>%
  filter(!is.na(estimate_log)) %>%
  group_by(scenario_no, replication_id, estimator) %>%
  mutate(
    estimate_hr = exp(estimate_log),
    is_worst_pred = estimate_hr == max(estimate_hr, na.rm = TRUE)
  ) %>%
  filter(is_worst_pred) %>%
  dplyr::select(scenario_no, replication_id, estimator, join_key) %>%
  rename(pred_worst_subgroup = join_key) %>%
  distinct()

# Compare predicted vs true worst subgroups
worst_accuracy <- predicted_worst %>%
  left_join(true_worst_subgroup, by = "scenario_no") %>%
  mutate(
    correctly_identified = pred_worst_subgroup == true_worst_subgroup
  ) %>%
  group_by(scenario_no, estimator) %>%
  summarise(
    n_replications = n(),
    n_correct = sum(correctly_identified, na.rm = TRUE),
    accuracy = n_correct / n_replications,
    .groups = 'drop'
  ) %>%
  arrange(scenario_no, estimator)

# Overall accuracy (across all scenarios)
worst_accuracy_overall <- predicted_worst %>%
  left_join(true_worst_subgroup, by = "scenario_no") %>%
  mutate(
    correctly_identified = pred_worst_subgroup == true_worst_subgroup
  ) %>%
  group_by(estimator) %>%
  summarise(
    n_replications = n(),
    n_correct = sum(correctly_identified, na.rm = TRUE),
    accuracy = n_correct / n_replications,
    .groups = 'drop'
  ) %>%
  mutate(
    estimator_display = beautify_rhs_estimator(estimator),
    accuracy_pct = getPercentage(accuracy, 1),
    accuracy_formatted = paste(n_correct, "/", n_replications, " (", accuracy_pct, ")", sep = "")
  ) %>%
  dplyr::select(estimator_display, accuracy_formatted) %>%
  arrange(estimator_display) %>%
  rename("RHS Model" = estimator_display, "Correct / Total" = accuracy_formatted)

cat("\nWorst Subgroup Identification Accuracy (Overall - across all scenarios):\n\n")
kbl(worst_accuracy_overall,
    caption = "RHS Methods: Accuracy of Identifying Worst Subgroup",
    booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE)
```

---

**Document generated**: `r Sys.time()`
