---
title: "SUNFISH Scenarios Generation"
output: html_document
---

# SUNFISH Trial Scenario Generation

This script generates simulated datasets for the SUNFISH Part 2 trial replication.

## Study Overview

**SUNFISH Part 2** was a randomized (2:1), double-blind, placebo-controlled trial:
- **N = 180** (120 Risdiplam : 60 Placebo)
- **Primary Endpoint**: Change from baseline in MFM32 total score at Month 12
- **Treatment Effect**: 1.55 points (Risdiplam 1.36 vs Placebo -0.19)
- **Standard Deviation**: 6.0 points
- **Key Finding**: Younger patients showed more pronounced improvements

**Covariates** (from Table 1 and Supplementary Appendix Figure S3):
- **x_1**: Age Group (2-5y [31%], 6-11y [32%], 12-17y [25%], 18-25y [12%])
- **x_2**: SMA Type (Type2 [71%], Type3 [29%])
- **x_3**: SMN2 Copy Number (2_copies [3%], 3_copies [87%], 4_copies [10%])
- **x_4**: Disease Severity based on Baseline MFM32 quartiles (Severe ≤Q1, Moderate, Mild >Q3)
- **baseline_mfm**: Baseline MFM32 score (continuous, Mean=46.11, SD=11.46)

**Covariate Correlations** (from SAP simulation documentation):
- Base correlation across all biological factors: ρ = 0.2
- Age × SMN2 Copy Number: ρ = 0.3 (Type 2 SMA tends to have fewer SMN2 copies)
- SMN2 Copy Number × Baseline MFM32: ρ = 0.35 (more copies → better motor function)
- Age × Baseline MFM32: ρ = -0.15 (older patients → more progression)

## Scenarios

We simulate **6 scenarios** adapted from the original bonsaiforest2 framework to match the SUNFISH trial context:

### 1. **Scenario 1 (Homogeneous)**: Primary efficacy analysis result
   - Consistent 1.55 point improvement across all subgroups
   - Represents the overall treatment effect reported in the trial

### 2. **Scenario 2 (Heterogeneous - Region-Driven)**: Dramatic regional heterogeneity from Forest Plot
   - **Europe (reference)**: +2.40 points (strong benefit)
   - **North America**: +2.44 points (similar to Europe)
   - **Japan**: -3.58 points (HARM - qualitative interaction)
   - **China**: -1.85 points (HARM - qualitative interaction)
   - **Rest of World**: +3.30 points (strong benefit)
   - Demonstrates **sign reversal**: treatment beneficial in Western regions, harmful in Asian regions
   - Most extreme form of treatment heterogeneity (qualitative interaction)

### 3. **Scenario 3 (Null/Crossover)**: Overall null with Type-based crossover
   - Overall mean treatment effect = 0
   - SMA Type 2 patients: +2.0 point improvement
   - SMA Type 3 patients: -2.0 point decline
   - Tests ability to detect crossover interactions

### 4. **Scenario 4 (Mild Random Heterogeneity)**: Small random noise
   - Base effect 1.55 points (overall population)
   - Random noise (SD=0.5) added to all subgroup interactions
   - Tests robustness to mild, unpredictable heterogeneity

### 5. **Scenario 5 (Large Random Heterogeneity)**: Large random noise
   - Base effect 1.55 points (overall population)
   - Large random noise (SD=2.0) added to all subgroup interactions
   - Tests robustness to strong, unpredictable heterogeneity

### 6. **Scenario 6 (Interaction)**: Age × SMA Type interaction
   - Base effect 2.5 points (young Type 2 patients)
   - Age reduces effect (-0.5 to -2.0 per age group)
   - Type 3 reduces effect (-0.8)
   - **Specific interaction**: Being both old (18-25y) AND Type 3 has additional penalty (-1.5)
   - Tests ability to detect complex 2-way interactions

## Library and Functions Loading

```{r setup}
# Ensure all required libraries are loaded
library(checkmate)
library(MASS) # Used by simul_covariates_sunfish

# --- LOAD SUNFISH FUNCTIONS ---
source('functions.R')

# --- REPRODUCIBILITY ---
RNGkind('Mersenne-Twister')
set.seed(0)
```

## Parameters Definition

```{r parameters}
# --- SIMULATION PARAMETERS ---
scenarios <- as.character(1:6)  # All 6 scenarios
n_datasets <- 1000              # Number of replicate datasets per scenario
```

## Data Generation Loop

```{r generate_data}
for (scen in scenarios) {
  
  # --- 1. Log Start ---
  message("------------------------------------------------------")
  message(paste("🚀 Starting: SUNFISH - Scenario", scen))
  
  # --- 2. Generate Data ---
  message("   ...generating data...")
  dataset_list <- simul_sunfish_data(
    scenario = scen,
    n_datasets = n_datasets
  )
  
  # --- 3. Define Save Path and Save ---
  save_path <- file.path("Scenarios", paste0("SUNFISH_Scenario_", scen, ".rds"))
  message(paste("   ...saving to", save_path))
  saveRDS(dataset_list, file = save_path)
  
  # --- 4. Log Completion ---
  message(paste("✅ Completed: SUNFISH - Scenario", scen))
  message(paste("   Saved", n_datasets, "datasets to", save_path))
  message("------------------------------------------------------")
  message("")
}
```

## Verification

```{r verify, eval=FALSE}
# Optional: Load and verify one dataset from each scenario
for (scen in scenarios) {
  message(paste("\n--- Verifying Scenario", scen, "---"))
  
  datasets <- readRDS(file.path("Scenarios", paste0("SUNFISH_Scenario_", scen, ".rds")))
  
  # Check first dataset
  dat <- datasets[[1]]
  
  message(paste("Sample size:", nrow(dat)))
  message(paste("Treatment allocation:"))
  print(table(dat$arm))
  message(paste("Age group distribution:"))
  print(table(dat$x_1))
  message(paste("SMA Type distribution:"))
  print(table(dat$x_2))
  message(paste("SMN2 Copy Number distribution:"))
  print(table(dat$x_3))
  message(paste("Disease Severity distribution:"))
  print(table(dat$x_4))
  message(paste("Baseline MFM32 - Mean:", round(mean(dat$baseline_mfm), 2), 
                "SD:", round(sd(dat$baseline_mfm), 2)))
  message(paste("Outcome (change) - Mean:", round(mean(dat$y), 2), 
                "SD:", round(sd(dat$y), 2)))
}
```

## Session Info

```{r session_info}
sessionInfo()
```
